### 2019-10-30 Cluster Timing/MB_PM_data.Rmd
Rearrange first step and spellcheck

---
title: "Mungbean Powdery Mildew Spray Timing"
author: "Adam H. Sparks"
date: "30/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(skimr)
mb <- read_csv("slim-mungbean-powdery_mildew.csv")
```

# Clean data

The csv file included is already reduced in columns compared with the original file.
This next step converts the date columns to `Date` class, removes more cols, including the dates for spray applications after two (there are up to seven in some studies) and creates two columns of time differences.

The first column of time differences is for the time from first sign to first spray. The second is for time from first sign to second spray.

Any rows that are missing yield data are removed.

```{r mutate-filter}
mb <-
  mb %>%
  select(-c(fungicide_application_3:fungicide_application_7)) %>%
  mutate(first_sign_disease = as.Date(first_sign_disease, format = "%d/%m/%Y")) %>%
  mutate_at(vars(starts_with("fungicide_application_")), dmy) %>%
  mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
  mutate(fungicide_timing_2 = fungicide_application_2 - first_sign_disease) %>%
  filter(!is.na(grain_yield.t.ha.))
```

# Inspect the new data set

```{r skim}
skim(mb)
```

# Create Histograms of time differences to spray

## First spray

```{r first-spray-hist}
ggplot(filter(mb, !is.na(fungicide_timing_1)),
       aes(x = as.numeric(fungicide_timing_1))) +
  geom_histogram(bins = 50, fill = "grey") +
  geom_vline(linetype = 2,
             xintercept = 0,
             colour = "red") +
  xlab("Days from first sign") +
  ggtitle("First Application",
          subtitle = "Line at first sign") +
  theme_bw()
```

Suggest that any sprays prior to first sign are a group with another group from Day 0 to Day 3 and third group being >= Day 4.

## Second spray

```{r second-spray-hist}
ggplot(filter(mb, !is.na(fungicide_timing_2)),
       aes(x = as.numeric(fungicide_timing_2))) +
   geom_histogram(bins = 50, fill = "grey") +
   geom_vline(linetype = 2,
              xintercept = 14,
              colour = "red") +
   xlab("Days from first sign") +
   ggtitle("Second Application",
           subtitle = "Line at 14 days after first sign") +
   theme_bw()
```

Again, suggest any sprays that are prior to 14 days after first sign be first group, the second being 14 days to 17 days after first sign and the third being >= 18 days after first sign.

These groups would then have preventative (before), at first sign and then 14 days after and late applications.

```{r cluster}
mb <-
  mb %>%
  mutate(
    cluster_1 = case_when(
      fungicide_timing_1 < 0 ~ "Early",
      fungicide_timing_1 >= 0 &
        fungicide_timing_1 < 3 ~ "Recommended",
      TRUE ~ "Late"
    )
  ) %>%
  mutate(
    cluster_2 = case_when(
      fungicide_timing_2 < 13 ~ "Early",
      fungicide_timing_2 >= 13 &
        fungicide_timing_2 < 17 ~ "Recommended",
      TRUE ~ "Late"
    )
  )
```

# Histograms with Clusters

## First spray

```{r hist-clust1}
ggplot(filter(mb, !is.na(fungicide_timing_1)),
       aes(x = as.numeric(fungicide_timing_1))) +
  geom_histogram(bins = 50, aes(fill = cluster_1)) +
  scale_fill_viridis_d("Cluster") +
   geom_vline(linetype = 2,
              xintercept = 0,
              colour = "red") +
   xlab("Days from first sign") +
   ggtitle("First Application",
           subtitle = "Line at first sign") +
   theme_bw()
```

## Second spray

```{r hist-clust2}
ggplot(filter(mb, !is.na(fungicide_timing_2)),
       aes(x = as.numeric(fungicide_timing_2))) +
  geom_histogram(bins = 50, aes(fill = cluster_2)) +
  scale_fill_viridis_d("Cluster") +
  geom_vline(linetype = 2,
             xintercept = 14,
             colour = "red") +
  xlab("Days from first sign") +
  ggtitle("Second Application",
          subtitle = "Line at 14 days after first sign") +
  theme_bw()
```

Note that the cluster for the second spray "recommended" is from Day 13, not Day 14.