### 2019-11-01 02_Weather.Rmd
Changed data import corrected some code from changed data files

# Weather data

The mungbean DSS determines the need to spray on:  
   1. The presence of the disease
   2. The forecasted maximum temperatures in the next seven days
   3. The number of days over the next seven days with rainfall

Below I am plotting the mean maximum temperatures over 7 days during the growing season at each mungbean trial site.
The vertical lines indicate when powdery mildew was first observed in that season.
Variation may exist in the sowing date of each trial.

```{r import_libraries_&_data}
source("R/import_data.R")
PM.dat <- import_data()  # importing means yield and disease severity for each treatment within each powdery mildew trial on mungbean
source("R/year2day.r")

```

```{r weather1, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# This chunk has been removed as the get_PPD2 function has stopped working and needs fixing
# Update: I have included data files instead of seeking the data from the web. 
# issue has something to do with website certificates not being appropriate

# Import dataset

# Need weather data between 2011 and 2017

# function to extract date from the year
# format should be Universal time format
source("R/year2day.r")

# Function to generate some of the ggplot grouping colours
gg_color_hue <- function(n) {
   hues = seq(15, 375, length = n + 1)
   hcl(h = hues, l = 65, c = 100)[1:n]
}
cols1 <- gg_color_hue(4)

silo_war_Tmax <- read.csv("./data/silo_war_Tmax.csv")
silo_king_Tmax <- read.csv("./data/silo_king_Tmax.csv")
silo_gatton_Tmax <- read.csv("./data/silo_gatton_Tmax.csv")
silo_emerald_Tmax <- read.csv("./data/silo_emerald_Tmax.csv")
```

### Kingaroy historical weather
```{r Kingaroy_weather, include=FALSE}
# Smoothing temperatures by taking a rolling7 day average
max_T <- vector()
min_T <- vector()
for (i in 1:length(silo_king_Tmax[, 1])) {
   max_T <- c(max_T, mean(silo_king_Tmax$max_temp[i:(i + 7)]))
   min_T <- c(min_T, mean(silo_king_Tmax$min_temp[i:(i + 7)]))
}

king_Tm_aver <-
   data.frame(
      location = "Kingaroy",
      date = as.Date(silo_king_Tmax$date),
      max_T,
      min_T,
      daily_rain = silo_king_Tmax$daily_rain
   )

# Add columns for Year and date to data
king_Tm_aver <- king_Tm_aver %>%
   mutate(
      Year = substr(king_Tm_aver$date, start = 1, stop = 4),
      Day = year2day(king_Tm_aver$date)
   )

# Create data subset of powdery mildew data means for the first sign of the disease by location
King_FS <- PM.dat %>%
   group_by(trial_ref, location, year) %>%
   summarise(first_sign = mean(as.Date(first_sign_disease), na.rm = T)) %>%
   filter(location == "Kingaroy") %>%
   mutate(day = year2day(first_sign))

ggplot(subset(king_Tm_aver, year == levels(factor(King_FS$year))),
       aes(day)) +
   geom_line(aes(y = max_T, colour = year)) +
   #geom_smooth(method = "lm", aes(y = max_T), colour = "darkred")+
   geom_line(aes(y = min_T, colour = Year)) +
   #geom_smooth(method = "lm", aes(y = min_T), colour = "darkblue")+
   labs(x = "Date", y = "Daily maximum temperature averaged over 7 days",
        title = "Maximum and minimum daily temperatures in Kingaroy") +
   scale_y_continuous(limits = c(0, 40)) +
   scale_x_continuous(
      limits = c(0, 200),
      labels = c(subset(king_Tm_aver, Year == 2016)$date[c(1, 50, 100, 150, 200)]),
      breaks = c(1, 50, 100, 150, 200)
   ) +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   geom_vline(
      xintercept = King_FS$day[1],
      size = 1,
      colour = cols1[1],
      linetype = "dotted"
   ) +
   geom_vline(
      xintercept = King_FS$day[2],
      size = 1.5,
      colour = cols1[2],
      linetype = "dotted"
   ) +
   geom_vline(
      xintercept = King_FS$Day[3],
      size = 1,
      colour = cols1[3],
      linetype = "dashed"
   ) +
   geom_area(aes(y = daily_rain, fill = Year), alpha = 0.5) +
   scale_y_continuous(sec.axis = dup_axis(name = "Rainfall (mm)"))
#   scale_x_continuous(limits = c(1,180),
#                      breaks = c(1,32,(31+29),(31+29+31),(31+29+31+30),(31+29+31+30+31),(31+29+31+30+31+30)),
#                      labels = c("2012-Jan-01","2012-Feb-01","2012-Mar-01","2012-Apr-01","2012-May-01","2012-Jun-01","2012-Jul-01"))
```

### Hermitage historical weather
```{r Hermitage_weather, include=FALSE}
# ______  Hermitage  _______
max_T <- vector()
min_T <- vector()
for (i in 1:length(silo_war_Tmax[, 1])) {
   max_T <- c(max_T, mean(silo_war_Tmax$max_temp[i:(i + 7)]))
   min_T <- c(min_T, mean(silo_war_Tmax$min_temp[i:(i + 7)]))
}

war_Tm_aver <-
   data.frame(
      location = "Hermitage",
      date = as.Date(silo_war_Tmax$date),
      max_T,
      min_T,
      daily_rain = silo_war_Tmax$daily_rain
   )

# Add columns for Year and date to data
war_Tm_aver <- war_Tm_aver %>%
   mutate(
      Year = substr(war_Tm_aver$date, start = 1, stop = 4),
      Day = year2day(war_Tm_aver$date)
   )

# Create data subset of powdery mildew data means for the first sign of the disease by location
Herm_FS <- PM.dat %>%
   group_by(Trial_ref, Location, Year) %>%
   summarise(First_sign = mean(as.Date(first_sign_disease), na.rm = T)) %>%
   filter(Location == "Hermitage") %>%
   mutate(Day = year2day(First_sign))

ggplot(subset(war_Tm_aver, Year == levels(factor(Herm_FS$Year))),
       aes(Day)) +
   geom_line(aes(y = max_T, colour = Year)) +
   #geom_smooth(method = "lm", aes(y = max_T), colour = "darkred")+
   geom_line(aes(y = min_T, color = Year)) +
   #geom_smooth(method = "lm", aes(y = min_T), colour = "darkblue")+
   labs(x = "Date", y = "Daily maximum temperature averaged over 7 days",
        title = "Maximum and minimum daily temperatures in Hermitage") +
   scale_y_continuous(limits = c(0, 45)) +
   scale_x_continuous(
      limits = c(0, 200),
      labels = c(subset(war_Tm_aver, Year == 2016)$date[c(1, 50, 100, 150, 200)]),
      breaks = c(1, 50, 100, 150, 200)
   ) +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   geom_vline(
      xintercept = Herm_FS$Day[1],
      size = 1,
      colour = cols1[1],
      linetype = "dotted"
   ) +
   geom_vline(
      xintercept = Herm_FS$Day[2],
      size = 1.5,
      colour = cols1[2],
      linetype = "dotted"
   ) +
   geom_vline(
      xintercept = Herm_FS$Day[3],
      size = 1,
      colour = cols1[3],
      linetype = "dashed"
   ) +
   geom_area(aes(y = daily_rain, fill = Year), alpha = 0.5) +
   scale_y_continuous(sec.axis = dup_axis(name = "Rainfall (mm)")) +
   scale_x_continuous(
      limits = c(1, 180),
      breaks = c(
         1,
         32,
         (31 + 29),
         (31 + 29 + 31),
         (31 + 29 + 31 + 30),
         (31 + 29 + 31 + 30 + 31),
         (31 + 29 + 31 + 30 + 31 + 30)
      ),
      labels = c(
         "Jan-01",
         "Feb-01",
         "Mar-01",
         "Apr-01",
         "May-01",
         "Jun-01",
         "Jul-01"
      )
   )

```

### Gatton historical weather
```{r Gatton_weather, include=FALSE}
# ______  Gatton  _______
max_T <- vector()
min_T <- vector()
for (i in 1:length(silo_gatton_Tmax[, 1])) {
   max_T <- c(max_T, mean(silo_gatton_Tmax$max_temp[i:(i + 7)]))
   min_T <- c(min_T, mean(silo_gatton_Tmax$min_temp[i:(i + 7)]))
}

# put maximum, minimum temperatures into a dataframe corresponding with their corresponding date
Gat_Tm_aver <-
   data.frame(
      location = "Gatton",
      date = as.Date(silo_gatton_Tmax$date),
      max_T,
      min_T,
      daily_rain = silo_gatton_Tmax$daily_rain
   )

# Add columns for Year and date to data
Gat_Tm_aver <- Gat_Tm_aver %>%
   mutate(
      Year = substr(Gat_Tm_aver$date, start = 1, stop = 4),
      Day = year2day(Gat_Tm_aver$date)
   )

# Create data subset of powdery mildew data means for the first sign of the disease by location
Gatton_FS <- PM.dat %>%
   group_by(Trial_ref, Location) %>%
   summarise(First_sign = mean(as.Date(first_sign_disease), na.rm = T)) %>%
   filter(Location == "Gatton") %>%
   mutate(Year = factor(substr(First_sign, 1, 4)),
          Day = year2day(First_sign))

ggplot(subset(Gat_Tm_aver, Year == 2012),
       aes(Day)) +
   geom_line(aes(y = max_T, colour = Year)) +
   #geom_smooth(method = "lm", aes(y = max_T), colour = "darkred")+
   geom_line(aes(y = min_T, colour = Year)) +
   #geom_smooth(method = "lm", aes(y = min_T), colour = "darkblue")+
   labs(x = "Date", y = "Daily maximum temperature averaged over 7 days",
        title = "Maximum and minimum daily temperatures in Gatton") +
   scale_y_continuous(limits = c(0, 40)) +
   scale_x_continuous(
      limits = c(0, 200),
      labels = c(subset(
         Gat_Tm_aver, Year == levels(Gatton_FS$Year)
      )$date[c(1, 50, 100, 150, 200)]),
      breaks = c(1, 50, 100, 150, 200)
   ) +
   theme(
      panel.grid.major = element_line(colour = "grey80"),
      panel.grid.minor = element_line(colour = "grey50"),
      panel.background = element_rect(fill = NA)
   ) +
   geom_vline(xintercept = Gatton_FS$Day,
              colour = "green",
              size = 1) +
   geom_area(aes(y = daily_rain, fill = Year), alpha = 0.5) +
   scale_y_continuous(sec.axis = dup_axis(name = "Rainfall (mm)")) +
   scale_x_continuous(
      limits = c(1, 180),
      breaks = c(
         1,
         32,
         (31 + 29),
         (31 + 29 + 31),
         (31 + 29 + 31 + 30),
         (31 + 29 + 31 + 30 + 31),
         (31 + 29 + 31 + 30 + 31 + 30)
      ),
      labels = c(
         "2012-Jan-01",
         "2012-Feb-01",
         "2012-Mar-01",
         "2012-Apr-01",
         "2012-May-01",
         "2012-Jun-01",
         "2012-Jul-01"
      )
   )
#   scale_x_continuous(limits = as.Date(c("2012-01-01","2012-06-01")))
```