### 2019-11-17 191115_PMMB_DataWrangling_PM.Rmd
Initialising Cleaning histortical data script into markdown file

# Data wrangling

## Import Libraries
```{r}
# install.packages("openxlsx")
# install.packages("readxl")
# install.packages("tidyr")
install.packages("agricolae")

library(openxlsx)
library(readxl)
library(tidyr)
library(dplyr)
library(data.table)
library(agricolae)

source("R/import_data.R")
PM_MB_means <- import_data()

```

## Functions
```{r functions}

fold_data <- function(new_DataFrame, template_DataFrame){
   for(i in colnames(new_DataFrame)){
      if(i == colnames(new_DataFrame)[1]){
         x1 <- vector()}

      cnames1 <- as.character(colnames(new_DataFrame))
      cnames2 <- as.character(colnames(template_DataFrame))

      x1 <- c(x1,match(i, cnames2, nomatch = NA)) # MAke a index vector for the location of new_DataFrame col names in template_DataFrame
   }

   new_DataFrame[cnames2[!(cnames2 %in% cnames1)]] <- NA # adding column names to new_DataFrame not in template_DataFrame
   new_DataFrame <- 
      new_DataFrame[,c(cnames2, cnames1[is.na(x1)])]  # adding column names to new_DataFrame not in template_DataFrame
   return(new_DataFrame)
}
```

## Data wrangling for each trial
### Hermitage 2010

```{r Hermitage_2010}

Herm_10 <-
   as.data.frame(read_xls(
      "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2010/2010 PMmung Herm/2010 PMmung hermitage RAW field book.xls",
      sheet = "For analysis"
   ), stringsAsFactors == FALSE)

Herm_10 <- Herm_10 %>%
   mutate(trade_name = case_when(`Schedule!` == "U/S" ~ NA_character_,
                                 `Schedule!` == "FirstsignC" ~ " ",
                                 TRUE ~ "Kendon"))%>%
   mutate(fungicide = case_when(`Schedule!` == "U/S" ~ "control",
                                 `Schedule!` == "FirstsignC" ~ "carbendazim",
                                 TRUE ~ "sulphur"))%>%
   mutate(dose_ai.ha = case_when(`Schedule!` == "U/S" ~ 0,
                                 `Schedule!` == "FirstsignC" ~ 250,
                                 TRUE ~ 2400))%>%
   mutate(fungicide_application_1 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-02-19", format = "%Y-%m-%d"),
                                              `Schedule!` == "Early Sp" ~ as.Date("2010-02-19", format = "%Y-%m-%d"),
                                              `Schedule!` == "FirstsignC" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
                                              `Schedule!` == "FirstsignS" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
                                              `Schedule!` == "U/S" ~ as.Date(NA),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_2 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-02-26", format = "%Y-%m-%d"),
                                              `Schedule!` == "Early Sp" ~ as.Date("2010-02-26", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_3 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-05", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_4 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-11", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_5 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-18", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_6 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-25", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_7 = case_when(`Schedule!` == "Full Sp" ~ as.Date("2010-03-07", format = "%Y-%m-%d"),
                                              TRUE ~ as.Date(NA)))%>%
   mutate(total_fungicide = case_when(`Schedule!` == "Full Sp" ~ 7,
                                      `Schedule!` == "Early Sp" ~ 2,
                                      `Schedule!` == "FirstsignC" ~ 1,
                                      `Schedule!` == "FirstsignS" ~ 1,
                                      `Schedule!` == "U/S" ~ 0,
                                      TRUE ~ 9999))

Herm_10_dat <- 
   data.frame(trial_ref = "mung0910/01",
              year = "2010",
              location = "Hermitage",
              host_genotype = Herm_10$`Cultivar!`,
              trial_design = "RB",
              plot_length = Herm_10$`Length (m)`,
              plot_width = 4,
              plant_density = NA,
              row_spacing = 0.25,
              replicate = Herm_10$`Rep!`,
              planting_date = as.Date("2010-01-22"),
              emergence_date = NA,
              flowering_date = NA,
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = as.Date("2010-03-18"),
              trade_name = Herm_10$trade_name,
              fungicide_ai = Herm_10$fungicide,
              dose_ai.ha = Herm_10$dose_ai.ha,
              n_treatment = Herm_10$`Treat!`,
              fungicide_application_1 = Herm_10$fungicide_application_1,
              fungicide_application_2 = Herm_10$fungicide_application_2,
              fungicide_application_3 = Herm_10$fungicide_application_3,
              fungicide_application_4 = Herm_10$fungicide_application_4,
              fungicide_application_5 = Herm_10$fungicide_application_5,
              fungicide_application_6 = Herm_10$fungicide_application_6,
              fungicide_application_7 = Herm_10$fungicide_application_7,
              total_fungicide = Herm_10$total_fungicide,
              harvest_date = as.Date("2010-05-14"),
              final_assessment = as.Date("2010-04-20"),
              PM_final_severity = Herm_10$P.M.13wap,
              rating_scale = "1-9",
              grain_yield.t.ha. = Herm_10$`Kg/ha`,
              AUDPC = apply(Herm_10[,c("P.M.9wap","P.M.11wap","P.M.13wap")],1, FUN = audpc, dates = c(9*7,11*7,13*7)),
              stringsAsFactors = FALSE
              )

Herm_10_m <- Herm_10_dat %>%
   group_by(trial_ref, year, location, host_genotype, trial_design, trade_name,fungicide_ai,dose_ai.ha,n_treatment) %>%
   summarise(plot_length.m. = mean(plot_length, na.rm = TRUE),
              plot_width.m. = mean(plot_width, na.rm = TRUE),
              plant_density = unique(plant_density),
              row_spacing = unique(row_spacing),
              replicates = sum(replicate),
              planting_date = unique(planting_date),
              emergence_date = NA,
              flowering_date = NA,
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = unique(first_sign_disease),
              fungicide_application_1 = unique(fungicide_application_1),
              fungicide_application_2 = unique(fungicide_application_2),
              fungicide_application_3 = unique(fungicide_application_3),
              fungicide_application_4 = unique(fungicide_application_4),
              fungicide_application_5 = unique(fungicide_application_5),
             fungicide_application_6 = unique(fungicide_application_6),
             fungicide_application_6 = unique(fungicide_application_7),
              total_fungicide = unique(total_fungicide),
              harvest_date = unique(harvest_date),
              final_assessment = unique(final_assessment),
              PM_final_severity = mean(PM_final_severity, na.rm = TRUE),
              rating_scale = unique(rating_scale),
              disease_error = sd(PM_final_severity, na.rm = TRUE),
              D_error_type = "stdev",
              grain_yield.t.ha. = mean(grain_yield.t.ha., na.rm = TRUE),
              raw_graded = NA,
              yield_error = sd(grain_yield.t.ha., na.rm = TRUE),
              Y_error_type = "stdev",
              AUDPC.m = mean(AUDPC),
             AUDPC.sd = sd(AUDPC),
              meanSq = anova(lm(grain_yield.t.ha. ~ host_genotype + factor(n_treatment), data = Herm_10_dat))[2,3]
   )

Herm_10_m <- fold_data(Herm_10_m, PM_MB_means)

write.csv(
   as.data.frame(Herm_10_m),
   "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2010/2010 PMmung Herm/2010 PMmung Hermitage means.csv"
)

```

### Hermitage 2011
```{r Herm_2011}

Herm_11 <-
   as.data.frame(read_xls(
      "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2011/2011 PMmung Herm/pmhermratings2011.xls",
      sheet = "Analyses factor"
   ), stringsAsFactors = FALSE)

Herm_11 <- Herm_11 %>%
   mutate(trade_name = 
             case_when(`Treatment!` == "1 spray AMI-X" ~ "Amistar Xtra",
                       `Treatment!` == "3 spray AMI-X" ~ "Amistar Xtra",
                       `Treatment!` == "1 spray SULPHUR" ~ "Kendon",
                       `Treatment!` == "3 spray SULPHUR" ~ "Kendon",
                       `Treatment!` == "1 spray FOLICUR" ~ "Folicur",
                       TRUE ~ "control"))%>%
   mutate(fungicide_ai = 
             case_when(`Treatment!` == "1 spray AMI-X" ~ "200 g/L azoxystrobin + 80 g/L cyproconazole",
                       `Treatment!` == "3 spray AMI-X" ~ "200 g/L azoxystrobin + 80 g/L cyproconazole",
                       `Treatment!` == "1 spray SULPHUR" ~ "sulphur",
                       `Treatment!` == "3 spray SULPHUR" ~ "sulphur",
                       `Treatment!` == "1 spray FOLICUR" ~ "tebuconazole",
                       TRUE ~ "control"))%>%
   mutate(dose_ai.ha = 
             case_when(`Treatment!` == "1 spray AMI-X" ~ 120,
                       `Treatment!` == "3 spray AMI-X" ~ 120,
                       `Treatment!` == "1 spray SULPHUR" ~ 3000,
                       `Treatment!` == "3 spray SULPHUR" ~ 3000,
                       `Treatment!` == "1 spray FOLICUR" ~ 62.35,
                       TRUE ~ 0))%>%
   mutate(fungicide_application_1 = 
             case_when(`Treatment!` == "1 spray AMI-X" ~ as.Date("2011-03-28", format = "%Y-%m-%d"),
                       `Treatment!` == "3 spray AMI-X" ~ as.Date("2011-02-28", format = "%Y-%m-%d"),
                       `Treatment!` == "1 spray SULPHUR" ~ as.Date("2011-03-28", format = "%Y-%m-%d"),
                       `Treatment!` == "3 spray SULPHUR" ~ as.Date("2011-02-28", format = "%Y-%m-%d"),
                       `Treatment!` == "1 spray FOLICUR" ~ as.Date("2011-03-28", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_2 = 
             case_when(`Treatment!` == "3 spray AMI-X" ~ as.Date("2011-03-14", format = "%Y-%m-%d"),
                       `Treatment!` == "3 spray SULPHUR" ~ as.Date("2011-03-14", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_3 = 
             case_when(`Treatment!` == "3 spray AMI-X" ~ as.Date("2011-03-14", format = "%Y-%m-%d"),
                       `Treatment!` == "3 spray SULPHUR" ~ as.Date("2011-03-14", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%
   mutate(total_fungicide = 
             case_when(`Treatment!` == "1 spray AMI-X" ~ 1,
                       `Treatment!` == "3 spray AMI-X" ~ 3,
                       `Treatment!` == "1 spray SULPHUR" ~ 1,
                       `Treatment!` == "3 spray SULPHUR" ~ 3,
                       `Treatment!` == "1 spray FOLICUR" ~ 1,
                       TRUE ~ 0))

Herm_11_dat <- 
   data.frame(trial_ref = "mung1011/01",
              year = "2011",
              location = "Hermitage",
              host_genotype = Herm_11$`Cultivar!`,
              trial_design = "RCB",
              plot_length = 11,
              plot_width = 4,
              plant_density = NA,
              row_spacing = 0.75,
              replicate = Herm_11$`Rep!`,
              planting_date = as.Date("2011-01-24", format = "%Y-%m-%d"),
              emergence_date = NA,
              flowering_date = as.Date("2011-03-22", format = "%Y-%m-%d"),
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = as.Date("2011-03-28", format = "%Y-%m-%d"),
              trade_name = as.factor(Herm_11$trade_name),
              fungicide_ai = Herm_11$fungicide_ai,
              dose_ai.ha = Herm_11$dose_ai.ha,
              n_treatment = Herm_11$Tr,
              fungicide_application_1 = Herm_11$fungicide_application_1,
              fungicide_application_2 = Herm_11$fungicide_application_2,
              fungicide_application_3 = Herm_11$fungicide_application_3,
              total_fungicide = Herm_11$total_fungicide,
              harvest_date = as.Date("2011-04-25", format = "%Y-%m-%d"),
              final_assessment = as.Date("2010-04-11", format = "%Y-%m-%d"),
              PM_final_severity = Herm_11$`P.M. 11  w.a.p.`,
              rating_scale = "1-9",
              grain_yield.t.ha. = Herm_11$`Yld Kg/ha`
              )

Herm_11_m <- Herm_11_dat %>%
   group_by(trial_ref, year, location, host_genotype, trial_design, trade_name,fungicide_ai,dose_ai.ha,n_treatment) %>%
   summarise(plot_length.m. = mean(plot_length, na.rm = TRUE),
              plot_width.m. = mean(plot_width, na.rm = TRUE),
              plant_density = unique(plant_density),
              row_spacing = unique(row_spacing),
              replicates = sum(replicate),
              planting_date = unique(planting_date),
              emergence_date = NA,
              flowering_date = unique(flowering_date),
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = unique(first_sign_disease),
              fungicide_application_1 = unique(fungicide_application_1),
              fungicide_application_2 = unique(fungicide_application_2),
              fungicide_application_3 = unique(fungicide_application_3),
              total_fungicide = unique(total_fungicide),
              harvest_date = unique(harvest_date),
              final_assessment = unique(final_assessment),
              PM_final_severity = mean(PM_final_severity, na.rm = TRUE),
              rating_scale = unique(rating_scale),
              disease_error = sd(PM_final_severity, na.rm = TRUE),
              D_error_type = "stdev",
              grain_yield.t.ha. = mean(grain_yield.t.ha., na.rm = TRUE),
              raw_graded = NA,
              yield_error = sd(grain_yield.t.ha., na.rm = TRUE),
              Y_error_type = "stdev",
              meanSq = anova(lm(grain_yield.t.ha. ~ host_genotype + factor(n_treatment), data = Herm_11_dat))[2,3]
   )

Herm_11_m <- fold_data(Herm_11_m, PM_MB_means)

write.csv(
   as.data.frame(Herm_11_m),
   "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2011/2011 PMmung Herm/2011 PMmung Herm means.csv"
)

```

### Kingaroy 2011
```{r King_11}
King_11 <-
   as.data.frame(read_xls(
      "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2011/2011 PMmung King/pmKing ratings2011.xls",
      sheet = "analyses"
   ), stringsAsFactors = FALSE)

tmp1 <- do.call(rbind.data.frame, strsplit(King_11$`Treatment!`, " "))
colnames(tmp1) <- c("cultivar" , "spray_n", "spray", "fungicide")
King_11 <- cbind(tmp1, King_11)

King_11 <- King_11 %>%
   mutate(trade_name = 
             case_when(fungicide == "AMIS-X" ~ "Amistar Xtra",
                       fungicide == "SULPHUR" ~ "Kendon",
                       fungicide == "FOLICUR" ~ "Folicur",
                       TRUE ~ "control"))%>%
   mutate(fungicide_ai = 
             case_when(fungicide == "AMIS-X" ~ "200 g/L azoxystrobin + 80 g/L cyproconazole",
                       fungicide == "SULPHUR" ~ "sulphur",
                       fungicide == "FOLICUR" ~ "tebuconazole",
                       TRUE ~ "control"))%>%
   mutate(dose_ai.ha = 
             case_when(fungicide == "AMIS-X" ~ 120,
                       fungicide == "SULPHUR" ~ 3000,
                       fungicide == "FOLICUR" ~ 62.35,
                       TRUE ~ 0))%>%
   mutate(fungicide_application_1 = 
             case_when(spray_n == 1 ~ as.Date("2011-03-23", format = "%Y-%m-%d"),
                       spray_n == 2 ~ as.Date("2011-03-23", format = "%Y-%m-%d"),
                       spray_n == 3 ~ as.Date("2011-03-09", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%

   mutate(fungicide_application_2 = 
             case_when(spray_n == 2 ~ as.Date("2011-04-13", format = "%Y-%m-%d"),
                       spray_n == 3 ~ as.Date("2011-03-30", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_3 = 
             case_when(spray_n == 3 ~ as.Date("2011-04-12", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))

King_11_dat <- 
   data.frame(trial_ref = "mung1011/02",
              year = "2011",
              location = "Kingaroy",
              host_genotype = King_11$cultivar,
              trial_design = "RCB",
              plot_length = 11,
              plot_width = 4,
              plant_density = NA,
              row_spacing = 0.75,
              replicate = King_11$`rep!`,
              planting_date = as.Date("2011-02-02", format = "%Y-%m-%d"),
              emergence_date = NA,
              flowering_date = NA,
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = as.Date("2011-03-22", format = "%Y-%m-%d"),
              trade_name = as.factor(King_11$trade_name),
              fungicide_ai = King_11$fungicide_ai,
              dose_ai.ha = King_11$dose_ai.ha,
              n_treatment = King_11$`trt!`,
              fungicide_application_1 = King_11$fungicide_application_1,
              fungicide_application_2 = King_11$fungicide_application_2,
              fungicide_application_3 = King_11$fungicide_application_3,
              total_fungicide = King_11$spray_n,
              harvest_date = as.Date(NA),
              final_assessment = as.Date("2010-05-11", format = "%Y-%m-%d"),
              PM_final_severity = King_11$`14 w.a.p.`,
              rating_scale = "1-9",
              grain_yield.t.ha. = as.numeric(King_11$`Yld kg ha`),
              AUDPC = apply(King_11[,c("9 w.a.p.", "11 w.a.p.", "12 w.a.p.", "14 w.a.p.")],1, FUN = audpc, dates = c(9*7,11*7,12*7, 14*7))
              )

King_11_m <- King_11_dat %>%
   group_by(trial_ref, year, location, host_genotype, trial_design, trade_name,fungicide_ai,dose_ai.ha,n_treatment) %>%
   summarise(plot_length.m. = mean(plot_length, na.rm = TRUE),
              plot_width.m. = mean(plot_width, na.rm = TRUE),
              plant_density = unique(plant_density),
              row_spacing = unique(row_spacing),
              replicates = sum(replicate),
              planting_date = unique(planting_date),
              emergence_date = NA,
              flowering_date = unique(flowering_date),
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = unique(first_sign_disease),
              fungicide_application_1 = unique(fungicide_application_1),
              fungicide_application_2 = unique(fungicide_application_2),
              fungicide_application_3 = unique(fungicide_application_3),
              total_fungicide = unique(total_fungicide),
              harvest_date = unique(harvest_date),
              final_assessment = unique(final_assessment),
              PM_final_severity = mean(PM_final_severity, na.rm = TRUE),
              rating_scale = unique(rating_scale),
              disease_error = sd(PM_final_severity, na.rm = TRUE),
              D_error_type = "stdev",
              grain_yield.t.ha. = mean(grain_yield.t.ha., na.rm = TRUE),
              raw_graded = NA,
              yield_error = sd(grain_yield.t.ha., na.rm = TRUE),
              Y_error_type = "stdev",
             AUDPC.m = mean(AUDPC),
             AUDPC.sd = sd(AUDPC),
              meanSq = anova(lm(grain_yield.t.ha. ~ factor(n_treatment), data = King_11_dat))[2,3]
   )

King_11_m <- fold_data(King_11_m, PM_MB_means)

write.csv(
   King_11_m,
   "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2011/2011 PMmung Herm/2011 PMmung Kingaroy means.csv"
)

```

### Goolhi 2013
```{r Goolhi_2013}

for(D_assessment in c("16 DAT1", "33DAT1")){

# Import final disease survey
Goolhi.13 <-
   as.data.frame(
      read_xlsx(
         path = "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2013/AM1305 Fungicides for powdery mildew in mungbean - Goolhi.xlsx",
         sheet = D_assessment,
         range = "A15:AG75",
         col_names = TRUE
      )
   )

Goolhi.13 <- Goolhi.13[order(Goolhi.13[,1],Goolhi.13[,2],Goolhi.13[,3],Goolhi.13[,4]),]

# Replace empty cells / NAs with zeros
for (i in 7:33) {
   Goolhi.13[is.na(Goolhi.13[, i]), i] <- 0
}

# Calculate incidence
#     Sites      1     2     3
# low canopy  c(7:9, 16:18, 25:27)
# mid canopy c(10:12, 19:21, 28:30)
# upp canopy c(13:15, 22:24, 31:33)

incidence1 <-
   vector(mode = "numeric", length = length(Goolhi.13[, 1]))

# Convert disease survey data to 1-9 Incidence rating scale
for (i in seq_along(incidence1)) {
   # incidence = 1
   if (all(Goolhi.13[i, 7:33] == 0)) {
      incidence1[i] <- 1
      next()
   }

   # incidence = 2
   if (all(Goolhi.13[i, c(10:15, 19:24, 28, 33)] == 0) &&
       (sum(Goolhi.13[i, c(7:9, 16:18, 25:27)] > 0) /
        length(Goolhi.13[i, c(7:9, 16:18, 25:27)])) <= 0.75) {
      incidence1[i] <- 2
      next()
   }

   # incidence = 3
   if (all(Goolhi.13[i, c(13:15, 22:24, 31:33)] == 0) &&
       sum(Goolhi.13[i, c(10:12, 19:21, 28:30)] != 0) <= 3 &&
       (sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] > 0) /
        length(Goolhi.13[i, c(7:12, 16:21, 25:30)])) > 0.75) {
      incidence1[i] <- 3
   } else{
      # if it is in the lower half of the canopy and in less than 75% of plants then incidence = 2.5
      if (all(Goolhi.13[i, c(13:15, 22:24, 31:33)] == 0) &&
          sum(Goolhi.13[i, c(10:12, 19:21, 28:30)] != 0) <= 3 &&
          (sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] > 0) /
           length(Goolhi.13[i, c(7:12, 16:21, 25:30)])) <= 0.75) {
         incidence1[i] <- 2.5
         next()
      }
   }

   # incidence = 4
   if (all(Goolhi.13[i, c(13:15, 22:24, 31:33)] == 0) &&
       sum(Goolhi.13[i, c(10:12, 19:21, 28:30)] != 0) > 3 &&
       (sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] > 0) /
        length(Goolhi.13[i, c(7:12, 16:21, 25:30)])) <= 0.75) {
      incidence1[i] <- 4
      next()
   }

   # incidence = 5
   if (all(Goolhi.13[i, c(13:15, 22:24, 31:33)] == 0) &&
       sum(Goolhi.13[i, c(10:12, 19:21, 28:30)] != 0) > 3 &&
       (sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] > 0) /
        length(Goolhi.13[i, c(7:12, 16:21, 25:30)])) > 0.75 &&
       any(Goolhi.13[i, c(7:12, 16:21, 25:30)] == 0)) {
      incidence1[i] <- 5
      next()
   }

   # incidence = 6
   if (all(Goolhi.13[i, c(13:15, 22:24, 31:33)] == 0) &&
       all(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0)) {
      incidence1[i] <- 6
      next()
   }

   # incidence = 7
   if (sum(Goolhi.13[i, c(13:15, 22:24, 31:33)] != 0) <= 3 &&
       all(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0)) {
      incidence1[i] <- 7
      next()
   }

   # incidence = 8
   if ((sum(Goolhi.13[i, c(13:15, 22:24, 31:33)] != 0) > 3 |
        sum(Goolhi.13[i, c(13:15, 22:24, 31:33)]) <= 200) &&
       all(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) &&
       (sum(Goolhi.13[i, 7:33] > 0) /
        length(Goolhi.13[i, 7:33]) > 0.75) &&
       sum(Goolhi.13[i, 7:33]) <= 2000) {
      incidence1[i] <- 8
      next()
   }

   # incidence = 9
   if (sum(Goolhi.13[i, c(13:15, 22:24, 31:33)] != 0) > 3 &&
       all(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) &&
       (sum(Goolhi.13[i, 7:33] > 0) /
        length(Goolhi.13[i, 7:33]) > 0.75) &&
       sum(Goolhi.13[i, 7:33]) > 2000) {
      incidence1[i] <- 9
      next()
   }

   # if not in the lower canopy but small colonies in mid and upper canopy
   if (all(Goolhi.13[i, c(7:9, 16:18, 25:27)] == 0) &&
       (sum(Goolhi.13[i, c(10:15, 19:24, 28, 33)] > 0) /
        length(Goolhi.13[i, 7:33]) <= 0.5)) {
      incidence1[i] <- 2.5
      next()
   }

   # incidence = 7.5
   if (sum(Goolhi.13[i, c(13:15, 22:24, 31:33)] != 0) > 3 &&
       sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) > 10 &&
       any(Goolhi.13[i, c(7:12, 16:21, 25:30)] == 0) &&
       (sum(Goolhi.13[i, 7:33] > 0) /
        length(Goolhi.13[i, 7:33]) > 0.7) &&
       sum(Goolhi.13[i, 7:33]) <= 2000) {
      incidence1[i] <- 7.5
      next()
   }

   # incidence = 6.5
   if (sum(Goolhi.13[i, c(13:15, 22:24, 31:33)] != 0) > 3 &&
       sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) <= 12 &&
       sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) >= 10 &&
       (sum(Goolhi.13[i, 7:33] > 0) /
        length(Goolhi.13[i, 7:33]) <= 0.75) &&
       sum(Goolhi.13[i, 7:33]) <= 2000) {
      incidence1[i] <- 6.5
      next()
   }

   # incidence = 4.5
   if (sum(Goolhi.13[i, c(13:15, 22:24, 31:33)] != 0) > 3 &&
       sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) < 10 &&
       sum(Goolhi.13[i, c(7:12, 16:21, 25:30)] != 0) > 5 &&
       (sum(Goolhi.13[i, 7:33] > 0) /
        length(Goolhi.13[i, 7:33]) <= 0.75) &&
       sum(Goolhi.13[i, 7:33]) <= 2000) {
      incidence1[i] <- 4.5
      next()
   } else{
      # If none of the conditions are met give the incidence zero
      incidence1[i] <- 0
   }

}

# check data to make sure there are no zeros
incidence1
length(Goolhi.13[incidence1 == 0, 1])
Goolhi.13[incidence1 == 0, ]
hist(incidence1)

# Re-format data-frame
if(D_assessment == "16 DAT1"){
Goolhi.13D <- Goolhi.13[,c("Treat", "Rep", "Run", "Plot")]
Goolhi.13D$Inc_16D <-  incidence1

}

if(D_assessment == "33DAT1"){
Goolhi.13D$Inc_33D <-  incidence1
}
}

# head(Goolhi.13D)
# Goolhi.13D <- Goolhi.13D %>%
#    group_by(Treat) %>%
#    summarise(M.inc = mean(Incidence, na.rm = TRUE),
#              sd.inc = sd(Incidence, na.rm = TRUE))

write.csv(
   Goolhi.13D,
   "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/Mungbean/Past trials/2013/AM1305-Goolhi-Disease_means.csv"
)

#_____________________
# Read in Harvest data
Goolhi.13Y <-
   as.data.frame(
      read_xlsx(
         path = "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/mungbean_powdery-mildew/Past trials/2013/AM1305 Fungicides for powdery mildew in mungbean - Goolhi.xlsx",
         sheet = "Yield",
         range = "A15:F75",
         col_names = TRUE
      )
   )

Goolhi.13Y <- Goolhi.13Y[order(Goolhi.13Y[,1],Goolhi.13Y[,2],Goolhi.13Y[,3],Goolhi.13Y[,4]),]

Goolhi_13 <- left_join(Goolhi.13D, Goolhi.13Y)

Goolhi_13 <- Goolhi_13 %>%
   mutate(trade_name = 
             case_when(Treat == "1" ~ NA_character_,
                       Treat == "2" ~ "Spin-flo",
                       Treat == "3" ~ "Tilt",
                       Treat == "4" ~ "Amistar Xtra",
                       Treat == "5" ~ "Amistar Xtra + adj",
                       Treat == "6" ~ "Cabrio",
                       Treat == "7" ~ "Spin-flo",
                       Treat == "8" ~ "Tilt",
                       Treat == "9" ~ "Amistar Xtra + adj",
                       Treat == "10" ~ "Cabrio",
                       Treat == "11" ~ "Tilt",
                       TRUE ~ "control"))%>%
   mutate(fungicide_ai = 
             case_when(fungicide == "AMIS-X" ~ "200 g/L azoxystrobin + 80 g/L cyproconazole",
                       fungicide == "SULPHUR" ~ "sulphur",
                       fungicide == "FOLICUR" ~ "tebuconazole",
                       TRUE ~ "control"))%>%
   mutate(dose_ai.ha = 
             case_when(fungicide == "AMIS-X" ~ 120,
                       fungicide == "SULPHUR" ~ 3000,
                       fungicide == "FOLICUR" ~ 62.35,
                       TRUE ~ 0))%>%
   mutate(fungicide_application_1 = 
             case_when(spray_n == 1 ~ as.Date("2011-03-23", format = "%Y-%m-%d"),
                       spray_n == 2 ~ as.Date("2011-03-23", format = "%Y-%m-%d"),
                       spray_n == 3 ~ as.Date("2011-03-09", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%

   mutate(fungicide_application_2 = 
             case_when(spray_n == 2 ~ as.Date("2011-04-13", format = "%Y-%m-%d"),
                       spray_n == 3 ~ as.Date("2011-03-30", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))%>%
   mutate(fungicide_application_3 = 
             case_when(spray_n == 3 ~ as.Date("2011-04-12", format = "%Y-%m-%d"),
                       TRUE ~ as.Date(NA)))

King_11_dat <- 
   data.frame(trial_ref = "mung1011/02",
              year = "2011",
              location = "Kingaroy",
              host_genotype = King_11$cultivar,
              trial_design = "RCB",
              plot_length = 11,
              plot_width = 4,
              plant_density = NA,
              row_spacing = 0.75,
              replicate = King_11$`rep!`,
              planting_date = as.Date("2011-02-02", format = "%Y-%m-%d"),
              emergence_date = NA,
              flowering_date = NA,
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = as.Date("2011-03-22", format = "%Y-%m-%d"),
              trade_name = as.factor(King_11$trade_name),
              fungicide_ai = King_11$fungicide_ai,
              dose_ai.ha = King_11$dose_ai.ha,
              n_treatment = King_11$`trt!`,
              fungicide_application_1 = King_11$fungicide_application_1,
              fungicide_application_2 = King_11$fungicide_application_2,
              fungicide_application_3 = King_11$fungicide_application_3,
              total_fungicide = King_11$spray_n,
              harvest_date = as.Date(NA),
              final_assessment = as.Date("2010-05-11", format = "%Y-%m-%d"),
              PM_final_severity = King_11$`14 w.a.p.`,
              rating_scale = "1-9",
              grain_yield.t.ha. = as.numeric(King_11$`Yld kg ha`),
              AUDPC = apply(King_11[,c("9 w.a.p.", "11 w.a.p.", "12 w.a.p.", "14 w.a.p.")],1, FUN = audpc, dates = c(9*7,11*7,12*7, 14*7))
              )

King_11_m <- King_11_dat %>%
   group_by(trial_ref, year, location, host_genotype, trial_design, trade_name,fungicide_ai,dose_ai.ha,n_treatment) %>%
   summarise(plot_length.m. = mean(plot_length, na.rm = TRUE),
              plot_width.m. = mean(plot_width, na.rm = TRUE),
              plant_density = unique(plant_density),
              row_spacing = unique(row_spacing),
              replicates = sum(replicate),
              planting_date = unique(planting_date),
              emergence_date = NA,
              flowering_date = unique(flowering_date),
              pod_fill_date = NA,
              mid_late_pod_fill = NA,
              first_sign_disease = unique(first_sign_disease),
              fungicide_application_1 = unique(fungicide_application_1),
              fungicide_application_2 = unique(fungicide_application_2),
              fungicide_application_3 = unique(fungicide_application_3),
              total_fungicide = unique(total_fungicide),
              harvest_date = unique(harvest_date),
              final_assessment = unique(final_assessment),
              PM_final_severity = mean(PM_final_severity, na.rm = TRUE),
              rating_scale = unique(rating_scale),
              disease_error = sd(PM_final_severity, na.rm = TRUE),
              D_error_type = "stdev",
              grain_yield.t.ha. = mean(grain_yield.t.ha., na.rm = TRUE),
              raw_graded = NA,
              yield_error = sd(grain_yield.t.ha., na.rm = TRUE),
              Y_error_type = "stdev",
             AUDPC.m = mean(AUDPC),
             AUDPC.sd = sd(AUDPC),
              meanSq = anova(lm(grain_yield.t.ha. ~ factor(n_treatment), data = King_11_dat))[2,3]
   )

King_11_m <- fold_data(King_11_m, PM_MB_means)

Goolhi.13Y <- Goolhi.13Y %>%
   group_by(Treat) %>%
   summarise(
      M.Yield = mean(`Yield Kg/ha`, na.rm = TRUE),
      sd.Yield = sd(`Yield Kg/ha`, na.rm = TRUE)
   )
write.csv(
   Goolhi.13Y,
   "C:/Users/U8011054/USQ/SCP - Documents/DAW1810/Mungbean/Past trials/2013/AM1305-Goolhi-Yield_means.csv"
)

```