### 2019-11-23 explore.Rmd
added netmeta results

---
title: "analysis"
author: "Emerson"
date: "10/29/2019"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(readxl)
library(janitor)
```

```{r}
datPM <- read_csv("data/slimmer_PM_clusterdat.csv")

```

# Fungicide treatments

```{r}
datPM %>% 
   tabyl(fungicide_ai)

```

```{r}

datPM %>%
   group_by(trial_ref, year, fungicide_ai) %>% 
   tally(sort = TRUE) %>% 
   ggplot(aes(x = fungicide_ai, fill = as.factor(year))) +
   geom_bar() +
   coord_flip()
```

```{r}
datPM %>% 
   filter(fungicide_ai %in% c("control", "tebuconazole", "propiconazole")) %>% 
   tabyl(year, fungicide_ai) %>% 
   gather(fungicide_ai, n, 2:4) %>% 
   ggplot(aes(factor(year), n, fill = fungicide_ai))+
   geom_col()
```

```{r}
datPM <- datPM %>% 
   mutate(V = case_when(is.na(Y_Msquare) ~ mean(datPM$Y_Msquare, na.rm = TRUE),
                    TRUE ~ Y_Msquare),
          yi = grain_yield.t.ha,
          vi = V/n, 
          id = 1:nrow(datPM))

```

```{r}
control <- datPM %>% 
   filter(spray_management == "control") %>% 
group_by(trial) %>% 
summarize(sev_control = mean(PM_final_severity))

control %>% 
   ggplot(aes(sev_control))+
   geom_histogram()

datPM2 <- left_join(control, datPM) %>% 
   filter(year != 2019)

datPM2 %>% 
   ggplot(aes(spray_management, yi, size = 1/vi, color = sev_control))+
   geom_jitter(width = 0.1)+
   facet_wrap(~year)

datPM2 %>% 
   filter(spray_management == "control") %>% 
   ggplot(aes(row_spacing, yi, size = 1/vi, color = sev_control))+
   geom_jitter(width = 0.1)

```

```{r}
library(metafor)

PM_mv1 <- rma.mv(yi, vi, 
                    mods = ~ spray_management, 
                    random = list(~ spray_management | trial, ~1 | id), 
                    struct="UN", data=datPM2)

```

```{r}
anova(PM_mv1, L=rbind(c(0,1,-1,0,0), 
                        c(0,0,1,-1,0),
                        c(0,0,1,0,-1),
                        c(0,0,0,-1,1),
                        c(0,1,0,0,1)))
```

```{r}
results <- data.frame(cbind(PM_mv1$b, 
                             PM_mv1$ci.lb,
                             PM_mv1$ci.ub))
results <- cbind(rownames(results), results)
rownames(results) <- NULL
colnames(results)<- c("treat","mean","lci", "uci")

results %>%
   filter(treat != "intrcpt") %>% 
   ggplot(aes(treat, mean))+
   geom_point(aes(size = 1/(uci-lci)), shape =15)+
   geom_errorbar(aes(ymin = lci, ymax = uci), width = 0.1)+
   coord_flip()

```

```{r}
library(netmeta)

datPM3 <- datPM2 %>% 
   group_by(trial, spray_management, n) %>% 
      summarize(yi_mean = mean(yi), 
             vi_mean = mean(vi)) %>% 
   ungroup()

PM_con <- pairwise(treat = spray_management,
                  n = n,
                  mean = yi_mean,
                  sd = sqrt(vi_mean),
                  studlab= trial,
                  data = datPM3, 
                  sm="MD")

```

```{r}
net_con <- netmeta(TE, 
                seTE, 
                treat1, 
                treat2, 
                studlab, 
                data=PM_con, 
                sm="MD")

summary(net_con)
```

```{r}
forest(net_con, reference.group = 4,
       rightcols=c("effect", "ci", "Pscore"),
       rightlabs="P-Score", small.values = "bad")
```

```{r}
netrank(net_con, small.values = "bad")
```

```{r}
netgraph(net_con, 
         plastic = FALSE,
         col = "orange",
         thickness =  "number.of.studies",
         points = FALSE, 
         col.points = "black", 
         cex.points = 1,
         number.of.studies = TRUE,
         cex.number.of.studies = 1,
         col.number.of.studies = "black",
         bg.number.of.studies = "orange",
         multiarm = FALSE,
         col.multiarm = "lightblue",
         pos.number.of.studies = 0.5)
```

```{r}
netleague(net_con)
```

```{r}
decomp.design(net_con)
```

```{r}
netsplit(net_con)
```
```{r}
nm1 <- netmeasures(net_con)

plot(nm1$meanpath, nm1$minpar, pch="",
     xlab="Mean path length", ylab="Minimal parallelism")
text(nm1$meanpath, nm1$minpar, names(nm1$meanpath), cex=0.8)
```