### 2020-01-16 05_MetaAnalysis.Rmd
More explaination to lmer

---
title: "Meta-analysis"
author: "P. Melloy"
date: "`r Sys.Date()`"
output: html_document
---

```{r MEsummary_Libraries, message=FALSE}
# install.packages("tidyverse")
# install.packages("devtools")
# devtools::install_github("https://github.com/PaulMelloy/bomrang")
# devtools::install_github("https://github.com/adamhsparks/theme.usq", dependencies = TRUE)
# install.packages("lme4")

library(RColorBrewer)
library(tidyverse)
library(theme.usq)
library(bomrang)
library(lme4)
library(kableExtra)
```

# Meta-analysis

```{r data_import, message=FALSE, echo=FALSE}
source("R/import_data.R")
PM_MB_means <- import_data()

# see Yield vs in-season rain section on the influence of rain and irrigation on grain yield
PM_MB_means$irrigation <- NA
PM_MB_means[PM_MB_means$location == "Hermitage", "irrigation"] <- TRUE
PM_MB_means[PM_MB_means$location != "Hermitage", "irrigation"] <- FALSE

```

## Estimating variance for trials without reported error

Now for the tricy part creating 'Sampling Variance' for treatments where only mean yield is recorded.  
The trial which does not report any type of variance also reports that there is no significant difference between the means.

```{r NAtrials}
NATrials <- PM_MB_means %>%
   filter(fungicide_ai == "tebuconazole",
          row_spacing == 0.75,
          is.na(Y_error_type)) %>%
   select(trial_ref) %>%
   distinct()
```

Two trials `mung1516/03` and `mung1112/01` within the scope of the meta-analysis fail to report variance statistics. According to [Machado F. et. al (2017)](http://apsjournals.apsnet.org/doi/10.1094/PDIS-03-17-0340-RE) [Paul et al. (2007)](https://apsjournals.apsnet.org/doi/pdfplus/10.1094/PHYTO-97-2-0211) we can use a linear regression model to estimate sampling variance.

## Preliminary meta-analysis

```{r LME_slimmer_cluster_disease_mod}
library(lme4)
slimmer_PM_dat <- read.csv("data/slimmer_PM_clusterdat.csv")

m8 <- lmer(grain_yield.t.ha*1000 ~ factor(spray_management) +
              (factor(spray_management) | trial_ref), 
           data = slimmer_PM_dat)

m9 <- lmer(log(grain_yield.t.ha*1000) ~ factor(spray_management) + 
              (factor(spray_management) | trial_ref), 
           data = slimmer_PM_dat)

m10 <- lmer(log(grain_yield.t.ha*1000) ~ 
               (factor(spray_management) | trial_ref), 
           data = slimmer_PM_dat)

m11 <- lmer(log(grain_yield.t.ha*1000) ~ factor(spray_management) + 
              (1 | trial_ref), 
           data = slimmer_PM_dat)

anova(m8,m9) # m9 is significantly better model
anova(m9,m10) # m9 is significantly better model
anova(m9,m11) # m11 is a simpler model which is no different to m9

summary(m9)

```

This linear mixed effect model shows indicates:  
   - A single early spray before first sign of powdery mildew is not likely to increase yields.  
   - A single or single spray with one or more follow sprays starting  at the `recommended` first spray , within 3 days of powdery mildew first sign are likely to produce significantly higher grain yields compared to the no spray control.  
   - The `recommended_plus` spray which has one or more follow-up sprays after first sign are likley to increase the mean grain yield.  
   - `Late_plus` spray treatments showed the highest mean grain yield, and was significantly higher than the no spray control. However showed no difference to either of the `recommended` treatments.

### Imputing sample variances

We need to impute the variances which are missing for a few of the trials. These trials `r unique(slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),c("location","year", "trial_ref")])` were analysed and reported that there was no significant difference between the treatments in each trial. So therefore we need the imputed variances not to show a significant difference.

There are `r sum(is.na(slimmer_PM_dat$yield_error))` treatments without yield error in our data, and `r sum(!is.na(slimmer_PM_dat$yield_error))` where yield error was recorded.

Plotting a histogram of the variances show that the yield is not normally distributed. A log transformation, however, shows a normal disrtubution. We can use the mean and standard deviation of the log(V) to sample variances for the treatments where V is missing.

```{r variance_histograms}
hist(slimmer_PM_dat$yield_error) # yield error is not normally distributed
hist(log(slimmer_PM_dat$yield_error)) # a log transformation shows it is normally distributed 
```

We will evaluate wheather the imputed variances confer a significant difference to the recorded means. I created a distance matrix of two standard errors above the mean against two standard errors below the mean. If any errors don't overlap, infering signicifant difference, the imputed variances are discarded and resampled until all treatment variances show no significant differences.

```{r imputing_variances}

# generate yield error values
for(i in 0:500){

   # Imputing missing log variances
   log_Yerror <- rnorm(n = sum(is.na(slimmer_PM_dat$yield_error)),
               mean = mean(log(slimmer_PM_dat$yield_error), na.rm = TRUE),
               sd = sd(log(slimmer_PM_dat$yield_error), na.rm = TRUE))

   # First lets convert Variance to standard error for all the variables that have variance recorded
   slimmer_PM_dat$yield_SE <- 
      sqrt(slimmer_PM_dat$yield_error)/sqrt(slimmer_PM_dat$n)

   # Second; adding the imputed standard errors
   slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error), "yield_SE"] <-
      sqrt(exp(log_Yerror))/sqrt(slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"n"])

# all errorbars should overlap so they are consistant with the meta-data describing no significant difference bettween treatments
# lets test to make sure
if(
   max(outer(X = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] -
               (2*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
             Y = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] +
               (2*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
             FUN = "-")
       ) > 0){

   message(paste("\nWarning!!: Imputed yield errors show significant differences when they should not, imputation will automatically rerun - iteration", "i =",i))
   next()
   }else{
      message(paste(i,"iterations: ", "imputed variances now show no significant distances :)\n Now adding variances to data "))

      # Plot the imputed errors for each trial
      Vplot <- slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),] %>%
         ggplot(aes(y = grain_yield.t.ha, trial))+
         #geom_point()+
         geom_pointrange(ymin = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] -
                            (1.95*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
                         ymax = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] +
                            (1.95*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
                         position = "jitter"
                         )+
         ylim(0.2,1.5)

      slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error), "yield_error"] <- exp(log_Yerror)
      print(Vplot)
      break()}
}

```

#### Imputing sample variances from Mean squares
```{r imputing_MSE}
slimmer_PM_dat <- read.csv("data/slimmer_PM_clusterdat.csv")
hist(unique(slimmer_PM_dat$Y_Msquare))
hist(log(unique(slimmer_PM_dat$Y_Msquare))) # log mean square has a more normal distribution

TrialMSQ <- slimmer_PM_dat %>%
   group_by(trial_ref)%>%
   summarise(unique(Y_Msquare))

TrialMSQ[is.na(TrialMSQ$`unique(Y_Msquare)`), "trial_ref"]

for(i in TrialMSQ[is.na(TrialMSQ$`unique(Y_Msquare)`),]$trial_ref){
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == i,"Y_Msquare"] <- 
      exp(
         rnorm(n = 1,
               mean = mean(log(TrialMSQ$`unique(Y_Msquare)`), na.rm = TRUE),
               sd(log(TrialMSQ$`unique(Y_Msquare)`), na.rm = TRUE)
         )
      )
   }

```

Before plotting lets have a look at how the trimmed down, modified data looks.

```{r treatment_means_plot}
slimmer_PM_dat$spray_management <- factor(slimmer_PM_dat$spray_management, levels(slimmer_PM_dat$spray_management)[rev(c(1,2,4,5,3))])

slimmer_PM_dat %>%
   ggplot(aes(y = prop_yield_gain, x = spray_management))+
   geom_boxplot()+
   #geom_point(position = "jitter", alpha = 1/5)+
   geom_jitter(width = 0.1, alpha = 1/5)+
   theme_usq()+
   labs(x = "Spray management variable", 
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield from each treatment \n categorised by spray management scenario")+
   theme(plot.title = element_text(hjust = 0.5))+
   coord_flip()

kableExtra::kable(table(slimmer_PM_dat$spray_management, slimmer_PM_dat$year), align=rep('c', 8))%>%  #commented as was redering empty, needs a look in
   kable_styling("striped", fixed_thead = TRUE, full_width = FALSE, position = "center")

```

## Meta-analysis
Let's load the metafor package we are using to analyse the data, then rearrange the factors we want to examine by placing the control treatment first. This way all treatments will be compared to the no spray controls.  
Next we are log transforming the grain yield and calculating the variance from the trial mean squares.  
Finally we will assign factor classes to the main variables in the meta-analysis.  
```{r meta_organisation}
library(metafor)
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management,c("control", "Early", "Recommended", "Recommended_plus","Late_plus"))

slimmer_PM_dat$yi <- log(slimmer_PM_dat$grain_yield.t.ha)
slimmer_PM_dat$vi <- slimmer_PM_dat$Y_Msquare/(slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha^2)

slimmer_PM_dat$spray_management <- factor(slimmer_PM_dat$spray_management)
slimmer_PM_dat$trial_ref <- factor(slimmer_PM_dat$trial_ref)
slimmer_PM_dat$trial <- factor(slimmer_PM_dat$trial)
```

```{r Meta-analysis}

PM_mv_AI <- rma.mv(yi,vi,
                   mods = ~ spray_management,
                   method = "ML",
                   random = list(~ spray_management | trial),
                   struct = "UN",
                   data = slimmer_PM_dat)

summary(PM_mv_AI)

```

```{r compare_meta&lmer}
lme_PM <- lmer(yi ~ spray_management + (spray_management | trial), data = slimmer_PM_dat)

summary(lme_PM)

```

```{r meta_contrasts}
anova(PM_mv_AI, L=rbind(c(0,1,-1,0,0),  # early vs late plus
                        c(0,1,0,-1,0),  # early vs recommended
                        c(0,1,0,0,-1),  # early vs recommended plus
                        c(0,0,-1,1,0),  # late plus vs recommended 
                        c(0,0,-1,0,1),  # late plus vs recommended  Plus
                        c(0,0,0,-1,1))) # recommended vs recommended plus
```
Results show that Early applications are not significantly different to the no spray control or the one and two spray management options at the recommended time. Lets view this on a plot.

```{r meta_results}
results_AI <- data.frame(cbind(exp(PM_mv_AI$b), 
                              exp((PM_mv_AI$ci.lb)),
                              exp(PM_mv_AI$ci.ub)))

results_AI <- data.frame(cbind(PM_mv_AI$b, 
                              (PM_mv_AI$ci.lb),
                              PM_mv_AI$ci.ub))

treat <- c("control", "Early", "Recommended", "Recommended_plus","Late_plus")
efficacy <- tbl_df(results_AI)
efficacy$treat <- treat
efficacy$se <- PM_mv_AI$se
colnames(efficacy) <- c("Mean", "CIsup", "CIinf", "Treatment", "SE")
efficacy
```

```{r meta_plot}
efficacy %>% 
  ggplot(aes(Treatment, Mean))+
  geom_point(aes(size=1/SE), shape=15)+
  geom_linerange(aes(ymin = CIinf, ymax = CIsup))+
  coord_flip()+
  theme_grey()+
   geom_hline(yintercept = efficacy[efficacy$Treatment == "control",]$CIinf)

```

Check the yeild

```{r}
boxplot(slimmer_PM_dat$grain_yield.t.ha ~ slimmer_PM_dat$spray_management)

```