### 2020-02-14 05_MetaAnalysis.Rmd
Adding more explaination to the analysis

```{r MEsummary_Libraries1, message=FALSE}
# install.packages("tidyverse")
# install.packages("devtools")
# devtools::install_github("https://github.com/PaulMelloy/bomrang")
# devtools::install_github("https://github.com/adamhsparks/theme.usq", dependencies = TRUE)
# install.packages("lme4")

library(RColorBrewer)
library(tidyverse)
library(theme.usq)
library(bomrang)
library(lme4)
library(kableExtra)

# Data
slimmer_PM_dat <- read.csv("data/slimmer_PM_clusterdat.csv")
```

# Meta-analysis
## Grain yield meta-analysis

```{r LME_slimmer_cluster_disease_mod}

m8 <- lmer(grain_yield.t.ha*1000 ~ factor(spray_management) +
              (factor(spray_management) | trial_ref), 
           data = slimmer_PM_dat)

m9 <- lmer(log(grain_yield.t.ha*1000) ~ factor(spray_management) + 
              (factor(spray_management) | trial_ref), 
           data = slimmer_PM_dat)

m10 <- lmer(log(grain_yield.t.ha*1000) ~ 
               (factor(spray_management) | trial_ref), 
           data = slimmer_PM_dat)

m11 <- lmer(log(grain_yield.t.ha*1000) ~ factor(spray_management) + 
              (1 | trial_ref), 
           data = slimmer_PM_dat)

anova(m8,m9) # m9 is significantly better model
anova(m9,m10) # m9 is significantly better model
anova(m9,m11) # m11 is a simpler model which is no different to m9

summary(m9)

```

This linear mixed effect model shows indicates:  
   - A single early spray before first sign of powdery mildew is not likely to increase yields.  
   - A single or single spray with one or more follow sprays starting  at the `recommended` first spray , within 3 days of powdery mildew first sign are likely to produce significantly higher grain yields compared to the no spray control.  
   - The `recommended_plus` spray which has one or more follow-up sprays after first sign are likley to increase the mean grain yield.  
   - `Late_plus` spray treatments showed the highest mean grain yield, and was significantly higher than the no spray control. However showed no difference to either of the `recommended` treatments.

### Imputing sample variances

We need to impute the variances which are missing for a few of the trials. These trials `r unique(slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),c("location","year", "trial_ref")])$trial_ref` were analysed and reported that there was no significant difference between the treatments in each trial. So therefore we need the imputed variances not to show a significant difference.

There are `r sum(is.na(slimmer_PM_dat$yield_error))` treatments without yield error in our data, and `r sum(!is.na(slimmer_PM_dat$yield_error))` where yield error was recorded.

Plotting a histogram of the variances show that the yield is not normally distributed. A log transformation, however, shows a normal disrtubution. We can use the mean and standard deviation of the log(V) to sample variances for the treatments where V is missing.

```{r variance_histograms}
hist(slimmer_PM_dat$yield_error) # yield error is not normally distributed
hist(log(slimmer_PM_dat$yield_error)) # a log transformation shows it is normally distributed 
```

We will evaluate wheather the imputed variances confer a significant difference to the recorded means. I created a distance matrix of two standard errors above the mean against two standard errors below the mean. If any errors don't overlap, infering signicifant difference, the imputed variances are discarded and resampled until all treatment variances show no significant differences.

```{r imputing_variances}

# generate yield error values
for(i in 0:500){

   # Imputing missing log variances
   log_Yerror <- rnorm(n = sum(is.na(slimmer_PM_dat$yield_error)),
               mean = mean(log(slimmer_PM_dat$yield_error), na.rm = TRUE),
               sd = sd(log(slimmer_PM_dat$yield_error), na.rm = TRUE))

   # First lets convert Variance to standard error for all the variables that have variance recorded
   slimmer_PM_dat$yield_SE <- 
      sqrt(slimmer_PM_dat$yield_error)/sqrt(slimmer_PM_dat$n)

   # Second; adding the imputed standard errors
   slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error), "yield_SE"] <-
      sqrt(exp(log_Yerror))/sqrt(slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"n"])

# all errorbars should overlap so they are consistant with the meta-data describing no significant difference bettween treatments
# lets test to make sure
if(
   max(outer(X = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] -
               (2*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
             Y = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] +
               (2*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
             FUN = "-")
       ) > 0){

   message(paste("\nWarning!!: Imputed yield errors show significant differences when they should not, imputation will automatically rerun - iteration", "i =",i))
   next()
   }else{
      message(paste(i,"iterations: ", "imputed variances now show no significant distances :)\n Now adding variances to data "))

      # Plot the imputed errors for each trial
      Vplot <- slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),] %>%
         ggplot(aes(y = grain_yield.t.ha, trial))+
         #geom_point()+
         geom_pointrange(ymin = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] -
                            (1.95*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
                         ymax = slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"grain_yield.t.ha"] +
                            (1.95*slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error),"yield_SE"]),
                         position = "jitter"
                         )+
         ylim(0.2,1.5)

      slimmer_PM_dat[is.na(slimmer_PM_dat$yield_error), "yield_error"] <- exp(log_Yerror)
      print(Vplot)
      break()}
}

```

The imputed variances calculated above don't show the expected positive correlation between grain yield and variance. This is unexpected and reflects the inherent variability of the data. Therefore we will calculate the trial mean squares and use this to calculate the variance for each sample.

#### Imputing sample variances from Mean squares
```{r imputing_MSE_dataSpread}
slimmer_PM_dat <- read.csv("data/slimmer_PM_clusterdat.csv")
hist(unique(slimmer_PM_dat$Y_Msquare))
hist(log(unique(slimmer_PM_dat$Y_Msquare))) # log mean square has a more normal distribution

```

Imputing using a log transformation of the data is required. Which trials need variance imputation?

```{r imputing_MSE}
TrialMSQ <- slimmer_PM_dat %>%
   group_by(trial_ref)%>%
   summarise(unique(Y_Msquare))

TrialMSQ[is.na(TrialMSQ$`unique(Y_Msquare)`), "trial_ref"]
```

```{r imputation_MSE}

for(i in TrialMSQ[is.na(TrialMSQ$`unique(Y_Msquare)`),]$trial_ref){
   slimmer_PM_dat[slimmer_PM_dat$trial_ref == i,"Y_Msquare"] <- 
      exp(
         rnorm(n = 1,
               mean = mean(log(TrialMSQ$`unique(Y_Msquare)`), na.rm = TRUE),
               sd(log(TrialMSQ$`unique(Y_Msquare)`), na.rm = TRUE)
         )
      )
   }

```

Before analysis lets have a look at the trimmed down modified data.

```{r treatment_means_plot}
slimmer_PM_dat$spray_management <- factor(slimmer_PM_dat$spray_management, levels(slimmer_PM_dat$spray_management)[rev(c(1,2,4,5,3))])

slimmer_PM_dat %>%
   ggplot(aes(y = grain_yield.t.ha, x = spray_management))+
   geom_boxplot()+
   #geom_point(position = "jitter", alpha = 1/5)+
   geom_jitter(width = 0.1, alpha = 1/5)+
   theme_usq()+
   labs(x = "Spray management variable", 
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield from each treatment \n categorised by spray management scenario")+
   theme(plot.title = element_text(hjust = 0.5))+
   geom_hline(yintercept = 0, linetype = 2)+
   coord_flip()
```

There seems like no difference between the treatments, with exception to Late_plus. Lets do this plot again, but use the proportion of yield increase compared to the no spray control as the response variable.

It is also important to visualise how well the data is compared across the trial years and trials.

```{r table_of_nTreatments}
kableExtra::kable(table(slimmer_PM_dat$spray_management, slimmer_PM_dat$year), align=rep('c', 8))%>%  #commented as was redering empty, needs a look in
   kable_styling("striped", fixed_thead = TRUE, full_width = FALSE, position = "center")

```
Treatments Late_plus and early don't have very good comparison to other treatments.

```{r plot_yieldProportion}
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management,rev(c("control", "Early", "Recommended", "Recommended_plus","Late_plus")))

slimmer_PM_dat %>%
   ggplot(aes(y = prop_yield_gain, x = spray_management))+
   geom_boxplot()+
   #geom_point(position = "jitter", alpha = 1/5)+
   geom_jitter(width = 0.1, alpha = 1/5)+
   theme_usq()+
   labs(x = "Spray management variable", 
        y = "Grain yield (t/Ha)",
        title = "Mean grain yield from each treatment \n categorised by spray management scenario")+
   theme(plot.title = element_text(hjust = 0.5))+
   geom_hline(yintercept = 0, linetype = 2)+
   coord_flip()
```

## Meta-analysis  
### metafor package  
Let's load the metafor package we are using to analyse the data, then rearrange the factors we want to examine by placing the control treatment first. This way all treatments will be compared to the no spray controls.  
Next we are log transforming the grain yield and calculating the variance from the trial mean squares.  
Finally we will assign factor classes to the main variables in the meta-analysis. Note that variable trial is a combination of:  
   - Trial identifier  
   - Trial year  
   - Trial location  
   - Host genotype  
   - Trial row spacing

```{r metafor_organisation}
library(metafor)
slimmer_PM_dat$spray_management <-
   factor(slimmer_PM_dat$spray_management,c("control", "Early", "Recommended", "Recommended_plus","Late_plus"))

slimmer_PM_dat$yi <- log(slimmer_PM_dat$grain_yield.t.ha)
slimmer_PM_dat$vi <- slimmer_PM_dat$Y_Msquare/(slimmer_PM_dat$n * slimmer_PM_dat$grain_yield.t.ha^2)

slimmer_PM_dat$spray_management <- factor(slimmer_PM_dat$spray_management)
slimmer_PM_dat$trial_ref <- factor(slimmer_PM_dat$trial_ref)
slimmer_PM_dat$trial <- factor(slimmer_PM_dat$trial)
```

```{r Metafor-analysis}

PM_mv_AI <- rma.mv(yi,vi,
                   mods = ~ spray_management,
                   method = "ML",
                   random = list(~ spray_management | trial),
                   struct = "UN",
                   data = slimmer_PM_dat)

summary(PM_mv_AI)

```

Results show that yields in single early spray treatments are not significantly different to the no spray control. Commencing spray manangement schedules at first sign of disease (Recommended), or between 7 - 19 days after first sign (late) produced significantly higher yields compared to the no spray control. On average a spray scheduale with two or more applications stating late (7-19 days after first sign of powdery mildew) produced the highest yields.

When we compare these results to those of the linear mixed effect model there is little difference in the outcome. The most noteable difference is the linear mixed-effect model is somewhat less certain about the `Late_plus` treatment, and the mean yield lower than the `recommended_plus` treatment.

```{r compare_meta&lmer}
lme_PM <- lmer(yi ~ spray_management + (spray_management | trial), data = slimmer_PM_dat)

summary(lme_PM)

```

To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.
```{r metafor_contrasts}
anova(PM_mv_AI, L=rbind(c(0,1,-1,0,0),  # early vs Recomended
                        c(0,1,0,-1,0),  # early vs recommended plus
                        c(0,1,0,0,-1),  # early vs Late plus
                        c(0,0,-1,1,0),  # Recommended vs recommended_plus
                        c(0,0,-1,0,1),  # Recommended vs Late_Plus
                        c(0,0,0,-1,1))) # recommended_plus vs Late_plus
```
Results show with the exclusion of the no spray control, none of the treatments are significantly different, however, early applications treatments on average produced lower yields. Lets view this on a plot.

```{r metafor_results}
results_AI <- data.frame(cbind(exp(PM_mv_AI$b), 
                              exp((PM_mv_AI$ci.lb)),
                              exp(PM_mv_AI$ci.ub)))

# results_AI <- data.frame(cbind(PM_mv_AI$b, 
#                               (PM_mv_AI$ci.lb),
#                               PM_mv_AI$ci.ub))

treat <- c("control", "Early", "Recommended", "Recommended_plus","Late_plus")
efficacy <- tbl_df(results_AI)
efficacy$treat <- treat
efficacy$se <- PM_mv_AI$se
colnames(efficacy) <- c("Mean", "CIs_lower", "CI_upper", "Treatment", "SE")
efficacy
```

```{r metafor_plot}
efficacy$Treatment <-
   factor(efficacy$Treatment,c("control", "Early", "Recommended", "Recommended_plus","Late_plus"))

efficacy %>% 
   ggplot(aes(Treatment, Mean))+
   geom_hline(yintercept = c(0.8,1,1.2,1.4), color = "grey80", linetype = 3)+
   geom_point(aes(size=1/SE), shape=15)+
   geom_linerange(aes(ymin = CIs_lower, ymax = CI_upper))+
   coord_flip()+
   theme_usq()

```

What is interesting here is that the variance in the control has increased while the treatments have decreased. This is despite the control being the best represented across all experiments. However it could also be because it also has a low number of pooled reps per trial. Lets look at how well each of the treatments compare to each other. We can use the netmeta package to give a graphical representation of this.

### netmeta package  
Lets analyse the data again using a different statistical approach to see if our outcome with the `metafor` package was robust. The `netmeta` package uses a frequentist approach to the analysis, and focuses on the pairwise comparisons between treatments.

```{r netmeta-analysis}
library(netmeta)

datPM3 <- slimmer_PM_dat %>% 
   group_by(trial, spray_management, n) %>% 
      summarize(yi_mean = mean(yi), 
             vi_mean = mean(vi)) %>% 
   ungroup()

PM_con <- pairwise(treat = spray_management,
                  n = n,
                  mean = yi_mean,
                  sd = sqrt(vi_mean),
                  studlab= trial,
                  data = datPM3, 
                  sm="MD")

net_con <- netmeta(TE, 
                seTE, 
                treat1, 
                treat2, 
                studlab, 
                data=PM_con, 
                sm="MD")

summary(net_con)

```
Now lets visualise this as a forest plot

```{r netmeta-forest}
forest(net_con, reference.group = 4,
       rightcols=c("effect", "ci", "Pscore"),
       rightlabs="P-Score", small.values = "bad")

```

The `netmeta` analysis suggests the spray scheduale commencing early are no different to any other treatment including the no spray control. It estimates the mean is very similar to the recommended treatments. The recommended plus and late_plus treatments show higher mean estimates, however not significantly different from the early estimate.

## AUPDC and PM severity meta-analysis
Now it's time to look at how the spray managment scenarios might impact disease pressure on the mungbean yields. First lets evaulate which measure of disease is the most robust predictor for yield loss, AUDPC, AUDPS or disease incidence at the end of the season.  
```{r LM_AUDPC}
library(MuMIn)
summary(lm(grain_yield.t.ha ~ AUDPC_m, data = slimmer_PM_dat))
summary(lm(grain_yield.t.ha ~ AUDPS_m, data = slimmer_PM_dat))
```

Area under the disease progress stairs (AUDPS) shows almost the same fit as the AUDPC linear model. AUDPC and AUDPS are not good predictors of loss in yield overall, without the inclusion of a `trial` parameter. Lets test the model with the inclusion of `trial` as a random effect.

```{r LMER_AUDPC1}
Me_mod1 <- lmer(grain_yield.t.ha ~ AUDPC_m + (1|trial),
              data = slimmer_PM_dat)
summary(Me_mod1)
r.squaredGLMM(Me_mod1)

Me_mod1.s <- lmer(grain_yield.t.ha ~ AUDPS_m + (1|trial),
              data = slimmer_PM_dat)
summary(Me_mod1.s)
r.squaredGLMM(Me_mod1)
r.squaredGLMM(Me_mod1.s)
anova(Me_mod1,Me_mod1.s)
```

AUDPC becomes a significant predictor of grain yield when the random effect when accounting for `trial`. However we noticed in the plots earlier that not only the intercept but the slope varies depending on the trial. Lets see if the model improves when AUDPC is used as a random slope for trial.

```{r LMER_AUDPC2}
Me_mod2 <- lmer(grain_yield.t.ha ~ AUDPC_m + (AUDPC_m|trial),
              data = slimmer_PM_dat)
anova(Me_mod1, Me_mod2)

```
The model is significantly improved, lower AIC and a higher log likelihood. Lets see if AUDPS improves in the same way.

```{r LMER_AUDPS2}
Me_mod2.s <- lmer(grain_yield.t.ha ~ AUDPS_m + (AUDPS_m|trial),
              data = slimmer_PM_dat)
anova(Me_mod1.s, Me_mod2.s)

```
The model did improve, however AUDPC still shows a better fit as shown by a lower AIC and a model which converged.

```{r AUDPCvsAUPDS}
r.squaredGLMM(Me_mod2.s) # Rsquared for AUDPStairs
r.squaredGLMM(Me_mod2)# Rsquared for AUDPCurve
```

The R-squared of the AUDPC model is a better fit compared to the AUDPS model.  
> From here on we will use AUDPC.

Lets compare the area under the disease progress curve (AUDPC) to final severity.
```{r LMER_finalSev}

Me_mod3 <- lmer(grain_yield.t.ha ~ PM_final_severity + (1|trial),
              data = slimmer_PM_dat)
Me_mod4 <- lmer(grain_yield.t.ha ~ PM_final_severity + (PM_final_severity|trial),
              data = slimmer_PM_dat)
anova(Me_mod3, Me_mod4)

summary(Me_mod4)
r.squaredGLMM(Me_mod4)

# PM_final_incidence vs AUDPC
anova(Me_mod4,Me_mod2)
```

Powdery mildew incidence at the end of the season was a significant covariate for grain yield, however was a weeker predictor compared to AUDPC. In addition model comparison shows AUDPC is significantly better at predicting grain yield.  
### Spray management and AUDPC
We shall proceed by including the spray management predictor into the analysis.  
First lets assess the `spray_management` variable as an additive fixed effect.

```{r AUDPC_spraymanagement}
Me_mod5 <- lmer(grain_yield.t.ha ~ AUDPC_m + spray_management+ (AUDPC_m|trial),
              data = slimmer_PM_dat)

anova(Me_mod2, Me_mod5)

```
This model is significantly better and shows a lower AIC. However makes more sense to use `spray_management` as a interactive term with AUDPC_m. As you would assume AUDPC would be dependant on the `spray_management`.

It is also worth noting there is no reason to include these parameters individually in the model to assess their independant effect on grain yield as we know the application of fungicide to mitigate yield loss is aimed at reducing the disease in the crop.

```{r AUDPC:spraymangement}
Me_mod6 <- lmer(grain_yield.t.ha ~ AUDPC_m : spray_management+ (AUDPC_m|trial),
              data = slimmer_PM_dat)

anova(Me_mod5, Me_mod6)
```

I will try to use the AUDPC:spray_management interative term as a random effect slope to trial, however I think this may overfit the model preventing convergance.

```{r AUDPC:spraymangement}
Me_mod8 <- lmer(grain_yield.t.ha ~ AUDPC_m : spray_management + (AUDPC_m: spray_management|trial),
              data = slimmer_PM_dat)

anova(Me_mod6, Me_mod8)
```
As expected the model produced an error warning stating the maximum number of evaluations were exceeded. Also the model comparisons show that the newer Me_mod7 was not a better model. We we proceed with Me_mod6 as the model with the best fit.

```{r LMER_AUDPC2_results}
summary(Me_mod6)
coef(Me_mod6)[["trial"]][
   order(coef(Me_mod6)[["trial"]]$AUDPC_m),]
```
The first output shows our random effects do a good job of explaining the variation leaving less residual variation than what is explained by the random intercept, trial.  
In the fixed effects the intercept gives an estimate of the average yield across all treatments (1.29 t/Ha). The other fixed effects show the change in predicted yield as AUDPC increases under each of the spray_management factors. Spray_management treatments, control, Early, and recommended show significant declines in grain_yields as AUDPC increases. Recommended_plus and Late_plus show no significant decline in grain yield as AUDPC increases.

The second output is the random effect coeficients. Here the intercept represents a yield estimate for each trial without the limitation of the disease. AUDPC_m shows the slope of the regression for grain yield as AUDPC increases. The coefficients of the model (ordered by AUDPC) show that higher yielding trials had a steeper negative slope. In other words higher yielding mungbean crops were likely to see greater losses from disease.

## AUPDC and PM severity meta-analysis