### 2020-04-03 06_Publication_wireframe_and_7_Questions.Rmd
writing and editing

---
title: Meta-analysis of fungicide application timing experiments for powdery mildew management
  in mungbean.
author: "P. Melloy"
date: "31/01/2020"
output:
  html_document: default
---

# Manuscript wire frame

## Who are the intended readers? - list 3 to 5 of them by name
   * Adam Sparks – Plant disease modeller
   * Levente Kiss – Powdery mildew taxonomist
   * Lisa Kelly – Plant pathologist
   * Paul MacIntosh – Mungbean/ pulse agronomist
   * Ross Pomroy – Agronomist
   * Jean Galloway – Plant disease modeller
   * Art Diggle – Plant disease modeller

## what did you do? (limit - 50 words)
Undertook a meta-analysis of twenty four field trials aimed at improving the efficiency and timing of fungicides for control of powdery mildew on mungbean.

## Why did you do it? (limit - 50 words)  
Data from unpublished field trials independently did not provide a clear indication for when to apply fungicide to best mitigate yield losses from powdery mildew. 
By analysing the collective data we aim to provide a better estimate for fungicide application timing to mitigate the yield loss from powdery mildew.

## what happened? (limit - 50 words)  
We found that fungicide sprays prior to when the powdery mildew was first spotted in the mungbean crop on average provided no yield protection. 
A fungicide spray schedule of one or more fungicide applications commencing within three days of first sign of the disease significantly mitigated yield loss to the disease. 
However a fungicide spray schedule of 2 or more applications commencing four days or later than first sign of disease saved the most yield.

## What do the results mean in theory? (limit - 50 words)
Fungicide spray schedules for control of powdery mildew in mungbean should commence around seven to nine days following the first signs of the disease.

## What do the results mean in practice? (limit - 50 words)  
Primarily that fungicide spray schedules that commence prior to signs of the disease establishing in the mungbean canopy are likely to be ineffective at mitigating the diseases impact on yield. 
We also don’t have data on at what point starting a fungicide spray schedule is too late. 
Growers can’t scout their crop three times a week in the fashion of researchers and are unlikely to identify the true date when signs of powdery mildew become apparent. 
Therefore these results reinforce the message to growers to spray for powdery mildew as soon as it is observed in their crop to protect yields.

## what is the key benefit for your readers? (limit - 25 words).  
This study is the most robust analysis of management of powdery mildew in mungbean using fungicides. 
It provides a more certain time frame for growers to plan when to commence a fungicide spray schedule.

Fewer fungicide applications will reduce costs for growers and prolong the efficacy of fungicides before the disease develops resistance.

## What remains unresolved? (no word limit).  
Within the meta-analysis, some comparisons between spray management factors contained fewer replicates in comparison to other spray management factor combinations.
Only a few studies incorporated in the meta-analysis contained comparisons between the “Late plus” factor and the treatments: negative spray control, a single spray at first sign, and spray at first sign with one or more follow-up applications.

*****

# Manuscript

## Introduction
Mungbean [*Vigna radiata* (L.) Wilczek] is a pulse crop primarily grown for human consumption in south-east Asia where it is an affordable source of protein [@PulseAustralia]. 
The bean pod or grain can be consumed raw or added to meals after sprouting. 
The dried grain can also be ground into a protein enriched flour for uses such as noodles, biscuits and cakes [@Chankaew2013].

In Australia mungbean is grown in southern Queensland and northern New South Wales as a short season summer legume crop. 
Approximately 90% of mungbean yields are exported providing good returns for growers. 
Mungbean potentially yields up to 2.3 tonnes per hectare, with a value of up to \$ 1300 a tonne [@QueenslandGovernment2019]. 
However these potential crop yields can be limited by a range of diseases[@Kelly2017a].
Powdery mildew (*Podosphaera xanthii*), for instance limits yields by up to 40% [@Chankaew2013]. 
Fortunately a number of strategies are available to be employed to reduce the damage powdery mildew inflicts on yields.

### Managements strategies
#### and conditions for infection
Planting date is an effective strategy for preventing the development of powdery mildew in mungbean. 
Powdery mildew infection requires cool temperatures, between 22$^\circ$ and 26$^\circ$C, and a leaf surface that is not overly wet for infection [@Kelly2017a]. 
In southern Queensland and northern New South Wales, sowing dates between late spring (November) and mid-summer (January) are advised to avoid the conducive cooler autumn temperatures which normally commence in March.  
Therefore a delayed sowing date increases the possibility powdery mildew will infect earlier in the crop development leading to greater yield losses.

Commercially available mungbean cultivars in Australia lack sufficient resistance to be used as a sole powdery mildew management strategy.
Quantitative disease resistance has been observed in some breeding lines overseas [@Pandey2018; @Chankaew2013]. 
While some quantitive resistance has also been incorporated into Australian varieties, only a couple of powdery mildew resistant cultivars are commercially available. 
Jade-AU is recommended as containing the best disease resistance to powdery mildew (moderately susceptible), while also confering quantifiable resistance to other foliar diseases such as tan spot and halo blight [@AustralianMungbeanAssociation]. 
Because cultivar choice does not confer complete resistance to powdery mildew additional strategies such as fungicide treatments are necessary to limit yield damage when conditions for the disease are suitable.

The recent release of 'PowderyMildewMDM' a desicion support system (DSS) increases the ease for preparing a fungicide spray scheduale to manage powdery mildew. 
PowderyMildewMDM provides a cost benefit analysis to growers for fungcide spray scheduale scenarios.
The user inputs weather, crop, and economic inputs into the DSS and the application returns the probability one or two fungicide sprays would protect yields suffiently to cover the costs of the fungicide applications and yield a profit.

### Fungicide resistance
Lucrative returns of \$ 2260 over a short three month season may tempt growers into applying fungicide to control diseases such as powdery mildew more than they are needed.

### Previous studies - fungicide work
There is almost a complete lack of peer-reviewed literature summarising fungicide efficacy on powdery mildew in Mungbean. 
Many of the experimental work has been published by funding agencies and state government extension departments. 
Past field trials in eastern Australia have trialed a range of fungicide active ingredients (table), while also attempting to evaluate the best application time for the highest efficacy [@goolhi2013; @premer2013; @Millmerran2013; @Marysmount2013; @SueThompson2016; @Kelly2017a; @Thompson2016]. 
Early trials showed for a fungicide application to provide maximum efficacy a single fungicide application at first sign of the disease with another follow-up application two weeks later if necessary [@Sparks2017]. 
Some of the trialled fungicides for treating powdery mildew in mungbean have been registered with the APVMA [@APVMA; @Kelly2017a; insert all published trials here].

### Meta-analysis

### Aims

*****

## Materials and methods

The data for this study was obtained from through personal correspondence with colleagues and collaborating institutions.
The majority of the trials were undertaken by Malcolm Riley and Dr Sue Thompson with The Department of Agriculture and Primary industries (DAF). 
Trials undertaken in 2013 season were conducted by the Northern growers alliance [@goolhi2013; @premer2013; @Millmerran2013; @Marysmount2013].
The most recent trials from 2017 were undertaken by DAF and the University of Southern Queensland (USQ) from 2017 to 2019 (Table 1). 
Trial data were in different formats with varying levels of information.
Some trial data reporting means of each treatment, with all but one study reporting variance along with the mean. 
For XX trials the raw data were obtained.

Details of trial data is presented within a research compendium which was prepared to supplement this publication [insert reference to compendium].

The criteria for inclusion of trial data in the meta-analysis, required a field trial testing fungicide efficacy for powdery mildew control on mungbean in the northern grains regions of eastern Australia.  
Trial data had to include the date when powdery mildew was first observed. 
Trials needed to have tested a demethylase inhibitor (DMI) fungicide.

All trials included in this analysis were not previously published in peer-reviewed literature.
Some trials and results were reported online via the Grains Research and Development Corporation.

While the main aim of all the trials collected for this study were the same, assess the efficacy of fungicides for control of powdery mildew, most also aimed to test the effect of different variables on fungicide efficacy. 
Some of the variable which were suspected to influence powdery mildew incidence and the effect on yield were: trial location, trial season, row spacing, fungicide active ingredient, fungicide dose, planting date, number of fungicide applications, cultivar, irrigation and rainfall.
To account for this variation between trials and thr treatments within the trial a single 'trial' factor was created to describe the main variables; Trial location, trial season, host genotype and row spacing.

To reduce confounding effects of various fungicides we only included treatments within trials which had been sprayed with a fungicide which has a DMI mode of action, such as tebeconazole and propiconazole. 
Fungicide active from the DMI group was the main inclusion factor for the analysis which reduced the number of trials from 24 to 17 experiments.

Most of the studies which were used in this analysis assessed powdery mildew incidence on a 1 - 9 ordinal scale [ref]. 
The 2013 NGA trials reported incidence as the percentage of leaves covered by powdery mildew and the location of the infected leaves in the canopy; lower, middle or upper canopy. 
Given the ordinal scale ranked plants on the percent of foliage up the plant showing the disease, it was easy to calculate the equivalent incidence on the ordinal 1 - 9 scale.

The timing of fungicide applications, with reference to when the disease first occurred in the trial, and number of applications were used to define a treatment variable.
An 'Early' fungicide application, refers to a fungicide spray schedule that commences with it's first spray prior to the disease being observed in the trial.  
'Recommended' refers to a fungicide spray schedule which commences within three days following the first sign of disease, and 'late' treatments which commence four days or later than first sign. 
The number of fungicide applications were binned in to two categories.
A spray schedule with a single spray, and a spray schedule with one or more sprays where there was a maximum of three applications. 
The division of the treatments into these six categories created two categories with insuffient numbers of treatments to provide a reliable sample. 
Single late sprays contained two replicates and mulitple early sprays contained five replicates and were therefore removed from the subsequent meta-analysis.

What remained was the mean yield for total of 173 treatments in the following five spray schedule categories: no spray control, single early spray, single recommended spray, recommended with multiple sprays and late with multiple sprays.

Trials where we obtained the raw data the response variables were summarised into the treatment means and variance. 
The remainder of the trial data reported treatment means and variance. 
Two of the trials, Gatton 2012 and Emerald 2015, reported treatment means without accompaning variance. 
For these trials the trial variance was imputed by randomly sampling the trial mean squared error from a distribution of log trial MSE.  
$rnorm(\overline{x}_{log(MSE_{1}) \dots, log(MSE_{n})}, \sigma_{log(MSE_{1}) \dots, log(MSE_{n})}) $  
Where MSE is the mean squared error of each trial.

$v_i = MSE_{y}/(n * Y_{i}^2) $

The data were analysed in the R statistical software environment using two meta-analysis packages, `metafor` and `netmeta`.

```{r Table1_ExperimentalDetail, echo=TRUE, message=FALSE, warning=FALSE}
PM.dat <- import_data()

PM.dat %>%
      group_by(trial_ref,) %>%
      summarise(year = unique(year),
                location = unique(location),
                planting_date = unique(planting_date),
                first_sign_disease = unique(first_sign_disease))%>%
      arrange(year)%>%
   left_join(data.frame(trial_ref = unique(PM.dat$trial_ref),
                        Research_org =
                           c(rep("Queensland DAF",times =12),
                             rep("Northern Growers Alliance", times = 4),
               rep("Queensland DAF",times =5),
               rep("USQ", times = 5))))%>%
kable(caption = "Table 1: Description of Experiments")%>%
   kable_styling(fixed_thead = TRUE,
                 full_width = TRUE) %>%
   scroll_box(width = "900px", height = "500px")
```