### 2020-05-22 01_Meta-analysis_experiments.Rmd
as per repository issue #20 column heading and contents made more clear and removed trial_ref

# Studies considered in meta-analysis

```{r MEsummary_Libraries, message=FALSE, echo=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, leaflet, htmltools,schoolmath)
source("R/import_data.R")
```

This meta-analysis uses data collected by a number of researchers from the Department of Agriculture and Primary Industries and the University of Southern
Queensland.
It is worth noting Professor Malcolm Riley's and Dr Sue Thompson's effort in establishing and coordinating early field trials between 2010 and 2016.

Studies were conducted in the following locations and years and were considered for inclusion in this analysis.

* 2001  
   + Bongeen #1  
   + Bongeen #2

* 2002  
   + Hermitage #1  
   + Hermitage #2  
   + Kingaroy

* 2004  
   + Redvale

* 2010  
   + Hermitage  
   + Kingaroy

* 2011  
   + Hermitage  
   + Kingaroy

* 2012  
   + Gatton  
   + Kingaroy

* 2013  
   + Premer  
   + Millmerran   
   + Marys Mount  
   + Goolhi

* 2015  
   + Hermitage  
   + Dalby - Site abandoned due to insufficient/late disease infection  
   + Emerald

* 2016  
   + Hermitage  
   + Kingaroy  
   + Emerald

* 2017  
   + Missen Flats (Fogerty's farm)  
   + Hermitage

* 2018  
   + Wellcamp

* 2019  
   + Hermitage

**2001 - 2004 trials**  
Trial data were obtained from a non-published manuscript.  
Ratings for powdery mildew were on a 1-5 rating scale where 1 = no sign of disease, 2 = 1-30% diseased leaf area, 3 = 31-65% diseased leaf area, 4 = 66-99% diseased leaf area and 5 = complete defoliation.  
Comparing trials which have not used the same rating scale will be done using a
percentage change.

**2013 Field trials**  
Trials were conducted by the [Northern Growers Alliance (NGA)](https://www.nga.org.au/).
The raw data were obtained from Lawrie Price.
NGA data were formatted to be compliant with existing data rating and scales using the R script [`R/Cleaning historical data.R`](https://github.com/openplantpathology/Mungbean_PM/blob/master/R/Cleaning%20historical%20data.R).

## Summary of data

```{r data_summary, echo=FALSE, message=FALSE, warning=FALSE}
PM.dat <- import_data()

kable(
   PM.dat %>%
      group_by(trial_ref, ) %>%
      summarise(
         year = unique(year),
         location = unique(location),
         replicates_per_treatment = if(is.whole(mean(replicates))){
            as.character(mean(replicates))}else{
               as.character(paste(min(replicates),"-",max(replicates)))},
         planting_date = unique(planting_date),
         first_sign_disease = unique(first_sign_disease),
         treatments = length(n_treatment)
      ) %>%
      arrange(year)%>%
      select(-trial_ref),
   caption = "Description of Experiments"
)  %>%
   kable_styling(fixed_thead = TRUE, full_width = TRUE) %>%
   scroll_box(width = "900px", height = "500px")

```

## Powdery mildew ratings

   1. No powdery mildew colonies observed on any plant

   2. Small colonies in lower 1/3 of canopy, up to 75% of plants affected

   3. Colonies in lower 1/2 of canopy, >75% of plants affected

   4. Colonies in lower 2/3 of canopy, up to 75% of plants affected

   5. Colonies in lower 2/3 of canopy, >75% of plants affected

   6. Colonies in lower 2/3 of canopy, 100% of plants affected

   7. Colonies in lower 2/3 of canopy of 100% of plants, some plants with colonies also in top 1/3 of canopy

   8. Colonies to top of plant with >75% of plants affected

   9. Colonies to top of plant with 100% of plants affected and heavy leaf drop

## Experiment locations

The following map displays experiment locations where data was sufficient to include in this meta-analysis.

```{r Experiment_locations, echo=FALSE}
experiment_sites <- data.frame(
   location = c(
      "Hermitage Research Station",
      "Red Vale & Kingaroy Research Station",
      "Gatton",
      "Emerald Agricultural College",
      "Bongeen 1 & 2",
      "NGA - Premer",
      "NGA - Millmerran",
      "NGA - Marys Mount",
      "NGA - Goolhi",
      "Missen Flats",
      "Wellcamp"
   ),
   lat = c(
      -28.215 ,
      -26.58,
      -27.56,
      -23.54,
      -27.51666,
      -31.43449,
      -27.923674,
      -31.0142,
      -31.00246,
      -27.905313,
      -27.564277
   ),
   lon = c(
      152.1003,
      151.83,
      152.32,
      148.19,
      151.450,
      150.0052,
      151.242706,
      150.05018,
      149.82507,
      151.9772,
      151.867879
   )
)

write.csv(experiment_sites,
          "cache/Mungbean_experiment_sites.csv",
          row.names = FALSE)

leaflet(experiment_sites) %>%
   addProviderTiles("Esri.WorldTopoMap") %>%
   setView(lat = -27.31,
           lng = 150.456,
           zoom = 5) %>%
   addMarkers(
      lng = ~ lon,
      lat = ~ lat,
      popup = ~ htmlEscape(location)
   )
```