### 2020-08-12 01_Meta-analysis_experiments.Rmd
Edit first chapter to follow the structure of the paper. Add the selection criteria and begin to implement it on the dataset

# Studies considered in meta-analysis

```{r MEsummary_Libraries, message=FALSE, echo=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, leaflet, htmltools, schoolmath, here)
source(here("R/import_data.R")
)
```
Trials testing fungicide efficacy on powdery mildew in mungbean in the Grains Research and Development Corporation (GRDC) northern grains region were sourced for a meta-analysis.

Data was collected from a number of researchers from the Department of Agriculture and Primary Industries and the University of Southern Queensland and the National Growers Association.
We would like to acknowledge in particular Professor Malcolm Riley's and Dr. Sue Thompson's effort in establishing and coordinating early field trials between 2010 and 2014.

Studies were collated from the following locations and years and were considered for inclusion in this analysis.

* 2001  
   + Bongeen #1  
   + Bongeen #2

* 2002  
   + Hermitage #1  
   + Hermitage #2  
   + Kingaroy

* 2004  
   + Redvale

* 2010  
   + Hermitage  
   + Kingaroy

* 2011  
   + Hermitage  
   + Kingaroy

* 2012  
   + Gatton  
   + Kingaroy

* 2013  
   + Premer  
   + Millmerran   
   + Marys Mount  
   + Goolhi

* 2015  
   + Hermitage  
   + Dalby - Site abandoned due to insufficient/late disease infection  
   + Emerald

* 2016  
   + Hermitage  
   + Kingaroy  
   + Emerald

* 2017  
   + Missen Flats (Fogerty's farm)  
   + Hermitage

* 2018  
   + Wellcamp

* 2019  
   + Hermitage

## Experiment locations

The following map displays experiment locations in the GRDC northern grain growing region which were considered for inclusion in this meta-analysis.

```{r Experiment_locations, echo=FALSE}
experiment_sites <- data.frame(
   location = c(
      "Hermitage Research Station",
      "Red Vale & Kingaroy Research Station",
      "Gatton",
      "Emerald Agricultural College",
      "Bongeen 1 & 2",
      "NGA - Premer",
      "NGA - Millmerran",
      "NGA - Marys Mount",
      "NGA - Goolhi",
      "Missen Flats",
      "Wellcamp"
   ),
   lat = c(
      -28.215 ,
      -26.58,
      -27.56,
      -23.54,
      -27.51666,
      -31.43449,
      -27.923674,
      -31.0142,
      -31.00246,
      -27.905313,
      -27.564277
   ),
   lon = c(
      152.1003,
      151.83,
      152.32,
      148.19,
      151.450,
      150.0052,
      151.242706,
      150.05018,
      149.82507,
      151.9772,
      151.867879
   )
)

write.csv(experiment_sites,
          "cache/Mungbean_experiment_sites.csv",
          row.names = FALSE)

leaflet(experiment_sites) %>%
   addProviderTiles("Esri.WorldTopoMap") %>%
   setView(lat = -27.31,
           lng = 150.456,
           zoom = 5) %>%
   addMarkers(
      lng = ~ lon,
      lat = ~ lat,
      popup = ~ htmlEscape(location)
   )
```

## Criteria for inclusion in meta-analysis

   1. A field trial testing the efficacy of fungicide on powdery mildew afflicted mungbean plants.  
   2. Trial data needed to include:  
      i) fungicide active ingredients.  
      ii) A Demethylase inhibitor (DMI) fungicide.  
      iii) The data at which powdery mildew first appeared in the trial.  
      iv) Disease severity at the end of the growing seasons.  
      v) Fungicide application dates.  
      vi) fungicide dose.  
      vii) Crop yield.  
      viii) Treatment means and accompanying variance

## Import data and apply selection criteria

### Import data from trials we have the raw data  
```{r import_dataRAW}
PM_MB_raw <- 
   rbind(
      read.csv(here("cache/2010 PMmung Hermitage means.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/2011 PMmung Herm means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/2011 PMmung Kingaroy means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1305_Goolhi_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1304-MarysMount_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1303-Premer-Disease_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/BB1305_Millmerran_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Herm_16_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/King_16_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Fogerty_17_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_17_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Wellcamp_18_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_19_mean.csv"),stringsAsFactors = FALSE)
      )
write.csv(PM_MB_raw, file = here("cache/PM_Mungbean_SummaryOfTrialsWithRawData"), row.names = FALSE)
```
### Import dataset of manual entries  
Import dataset which were manually entered from trial reports
```{r import_dataMan}
source(here("R/import_ManData.R"))
PM_MB_man <- import_ManData()
```

Bind datasets, data set compiled from raw trial data and data manually compiled from trial reports.
```{r bind_data}
PM_MB_dat <- 
   rbind(
      PM_MB_raw,
      PM_MB_man
      )
```

## Summary of all trials  
```{r data_summary, echo=FALSE, message=FALSE, warning=FALSE}
PM_MB_dat %>%
   arrange(year, location) %>%
   group_by(trial_ref) %>%
   summarise(
         Year = unique(year),
         Location = unique(location),
         `Replicates per treatment` = if(all(replicates[1] == replicates)){
            as.character(replicates[1])}else{
               as.character(paste(min(replicates),"-",max(replicates)))},
         `Planting date` = unique(planting_date),
         `First sign of disease` = unique(first_sign_disease),
         `Fungicide treatments` = length(n_treatment),
      ) %>%
      arrange(Year)%>%
   kable(caption = "Description of Experiments",
         align = "c",
         col.names = c("Unique Trial\nReference","Year", "Location", "Replicates\nper treatments",
                       "Planting date", "First sign\nof disease", "Fungicide\n treatments")
         )  %>%
   kable_styling(fixed_thead = TRUE, full_width = TRUE) %>%
   column_spec(c(3,5:6), width = "3cm") %>% 
   scroll_box(height = "500px")

```

## Powdery mildew rating scales

**2001 - 2004 trials**

Trial data were obtained from a non-published manuscript.

### 1 - 5 powdery mildew rating scale

Ratings for powdery mildew were on a 1 - 5 rating scale where

   1 - No sign of disease

   2 - 1 - 30&nbsp;% diseased leaf area

   3 - 31 - 65&nbsp;% diseased leaf area

   4 - 66 - 99&nbsp;% diseased leaf area and

   5 - Complete defoliation

Comparing these trials which have not used the 1 - 9 rating scale will be converted using a percentage change.

**2013 Field trials**

Trials from 2013 were all conducted by the [Northern Growers Alliance (NGA)](https://www.nga.org.au/).
The raw data were obtained from Lawrie Price.
NGA data were formatted to be compliant with the 1 - 9 rating scale used in all other experiments.
Detailed code for the conversion can be found in the following vignette [191115_PMMB_DataWrangling_PM.Rmd](./vignette/191115_PMMB_DataWrangling_PM.html).

### 1 - 9 powdery mildew rating scale

   1 - No powdery mildew colonies observed on any plant

   2 - Small colonies in lower 1/3 of canopy, up to 75&nbsp;% of plants affected

   3 - Colonies in lower 1/2 of canopy, >75&nbsp;% of plants affected

   4 - Colonies in lower 2/3 of canopy, up to 7&nbsp;5% of plants affected

   5 - Colonies in lower 2/3 of canopy, >75&nbsp;% of plants affected

   6 - Colonies in lower 2/3 of canopy, 100&nbsp;% of plants affected

   7 - Colonies in lower 2/3 of canopy of 10&nbsp;0% of plants, some plants with colonies also in top 1/3 of canopy

   8 - Colonies to top of plant with >75&nbsp;% of plants affected

   9 - Colonies to top of plant with 100&nbsp;% of plants affected and heavy leaf drop