### 2020-08-12 03_Preliminary_analysis.Rmd
Update preliminart analysis page

# Preliminary Analysis

```{r PA_Libraries, include=FALSE, message=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               lubridate,
               bomrang,
               openxlsx,
               devtools,
               reshape2,
               lme4,
               kableExtra,
               here)

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq()
)
load("cache/ImportDataAndSelectTrials01.Rdata")
```

## Explore and Visualise Data
Each row provides the mean yield for each experimental treatment.

The data explored in the `PM_MB_dat` data frame is a list of mean values from each treatment in each respective experiment.
These preliminary analyses will focus on the change in mungbean yields, which is attributable to the fungicide treatment. 
Later in this document we will also consider the impact of fungicide treatments on disease incidence/severity.
Mean powdery mildew incidence was recorded at the end of each growing season.
Some experiments recorded disease incidence throughout the season and these were summarised with AUDPC.
AUDPC for trials which only provided final powdery mildew incidence was calculated using the record of 'first sign of disease' and final assessment.

Various factors have been studied in the previous experiments which may influence the subsequent meta-analysis.
On this page, unless stated, the following plots and analyses were made disregarding other effects.
The following factors are postulated to influence mungbean grain yield and/or powdery mildew mean plot severity.

   1 Fungicide type

   2 Fungicide dose (varies only slightly within some fungicide types)

   3 Number of fungicide sprays

   4 Timing of fungicide spray/s relative to first sign of disease

   5 Host cultivar (probably is a co-variate with season due to changing cultivars over time)

   6 Experiment location

   7 Row spacing

   8 Final disease rating

The data plotted below are from `r length(unique(PM_MB_dat$trial_ref))` field trials between (`r first(levels(PM_MB_dat$year))` - `r last(levels(PM_MB_dat$year))`) of which the details are described in the [Trials considered for inclusion in meta-analysis section](#Trials-considered-for-inclusion-in-meta-analysis).

### Fungicides

#### Fungicide type

Before we exclude non-DMI fungicides let's have a look at the efficacy of the other fungicides in comparison to the DMIs.
For this comparison we will focus on the proportional yield saved, relative to the no spray control.
The primary focus of this research is the effect of fungicide on mitigating yield loss due to disease.
We are using proportional yield, relative to the no spray control, to reduce the effect of location and seasonal influences.

Let's investigate the difference in the effect of the type of fungicide on proportion of yield saved, but first we need to calculate yield saved relative to the no spray control.

```{r Calculate_yield-savings}
for (i in unique(PM_MB_dat$trial_ref)) {
   # Loop code over each trial reference
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]

   if (any(is.na(unique(dat1$row_spacing)))) {
      # If there is no recorded row.spacing don't calculate proportional yield, skip to the next trial
      warning(
         unique(dat1$trial_ref),
         "at",
         unique(dat1$location),
         "in year",
         unique(dat1$year),
         "has unknown row.spacing (NA)\n 'yield_gain' and 'proportional yield gain' not calculated\n"
      )
      next()
   }

   for (j in unique(dat1$row_spacing)) {
      # Loop within trial row_spacing
      dat2 <- dat1[dat1$row_spacing == j, ]
      controlY <-
         mean(dat2[dat2$fungicide_ai == "control", "grain_yield.t.ha."]) #Mean taken as some trials had multiple control reps

      # Calculate the yield gain relative to the no spray control (controlY)
      dat2$yield_gain <- dat2$grain_yield.t.ha. - controlY
      dat2$prop_YG <- dat2$yield_gain  /  controlY

      # insert NA as yeild_gain and proportional yield gain for the controls
      dat2[dat2$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
         NA

      # Reassign values to row_spacing subset
      dat1[dat1$row_spacing == j, ] <- dat2
   }
   # Reassign values to trial_ref subset
   PM_MB_dat[PM_MB_dat$trial_ref == i, ] <- dat1
}

# ____________________________________________
# calculate yield gain and prop_yield gain for experiments with no row spacing data
dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == "AM1303", ]

controlY <-
   mean(dat1[dat1$fungicide_ai == "control", "grain_yield.t.ha."])

dat1$yield_gain <- dat1$grain_yield.t.ha. - controlY
dat1$prop_YG <- dat1$yield_gain / controlY
dat1[dat1$fungicide_ai == "control", c("yield_gain", "prop_YG")] <-
   c(NA, NA)

PM_MB_dat[PM_MB_dat$trial_ref == "AM1303", ] <- dat1
```

Now we have calculated proportional yield savings we can inspect the difference between the fungicide active ingredients

```{r Fungicide_x_nSprays_tble2, message=FALSE}
PM_MB_dat %>%
   group_by(fungicide_ai, total_fungicide) %>%
   summarise(
      n = length(prop_YG),
      lower_2.5 = quantile(prop_YG, na.rm = T, c(0.025)),
      median = median(prop_YG, na.rm = T),
      mean = mean(prop_YG, na.rm = T),
      upper_97.5 = quantile(prop_YG, na.rm = T, c(0.975))
   ) %>%
   filter(n >= 5) %>% # remove any fungicide groups with less than 5 observations
   arrange(desc(median)) %>%
   kable(
      caption = "Fungicide effect on proportion of yield gain.",
      col.names = c(
         "Fungicide active ingredient",
         "Total fungicide applications",
         "n Treatments",
         "2.5% quartile",
         "Median proportional grain yield",
         "Mean proportional grain yield",
         "97.5 % quartile"
      ),
      align = "c"
   ) %>%
   kable_styling(fixed_thead = T, full_width = T) %>%
   column_spec(c(2, 3), width = "1cm") %>%
   footnote(general = "Fungicides with less than five observations were omitted from this table")
```

*****

Let's visualise this in a plot for one and two sprays only.

```{r Fungicide_x_nSprays_tble1, message=FALSE, warning=FALSE}
PM_MB_dat %>%
   mutate(fungicide_ai = as.factor(PM_MB_dat$fungicide_ai)) %>%
   filter(total_fungicide == 1 | total_fungicide == 2) %>%
   filter(
      fungicide_ai != "pyrazophos" &
         fungicide_ai != "control" &
         fungicide_ai != "benomyl" &
         fungicide_ai != "Acibenzolar-S-methyl"
   ) %>%
   mutate(fungicide_ai = factor(
      fungicide_ai,
      levels = c(
         "carbendazim",
         "pyraclostrobin",
         "sulphur",
         "200 g/L azoxystrobin + 80 g/L cyproconazole",
         "tebuconazole",
         "propiconazole",
         "200 g/L tebuconazole + 120 g/L azoxystrobin"
      )
   )) %>%
   ggplot(aes(y = prop_YG, x = fungicide_ai)) +
   facet_grid(rows = vars(total_fungicide)) +
   geom_boxplot(aes(fill = fungicide_ai)) +
   geom_hline(aes(yintercept = 0), size = 0.5) +
   labs(y = "Proportion of grain yield saved",
        title = "Grain yield proportion saved\nfaceted by number of sprays") +
   theme(legend.position = "none") +
   scale_fill_usq() +
   coord_flip()
```

There is a small yield effect among the strobilurin fungicide mixes and the DMI fungicides. 
While there is some differences in the variation between tebuconazole and propiconazole, the medians are very similar. 
I think it is safe to combine these fungicides, with the same mode of action in the meta-analysis.
The follow-up spray seems to show an increase in yield protection.

*****

#### Fungicide Doses  
From this point on we will investigate only the DMI fungicides, tebuconazole and propiconazole, and disregard all other fungicide modes-of-action.

We should check that all fungicide doses that were used were roughly the same, if we are to compare between trials where dose might be different.

```{r tebuAndPropi_dose}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   select(
      trial_ref,
      year,
      location,
      first_sign_disease,
      fungicide_ai,
      dose_ai.ha,
      total_fungicide
   ) %>%
   ggplot(aes(x = as.factor(dose_ai.ha), fill = fungicide_ai)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Total number of treatments for each respective tebuconazole\nor propiconazole dose") +
   geom_bar() +
   scale_fill_usq() +
   scale_colour_usq()
```

All trials that used tebuconazole used approximately the same dose.
Dose of the active ingredient ranged from 62.35&nbsp;g per hectare to 60&nbsp;g per hectare.

Doses for propiconazole range from 62.5 to 125, a large variation.
Let's inspect the difference in yields for each dose.

```{r PropiDoseEffect}
PM_MB_dat %>%
   filter(fungicide_ai == "propiconazole") %>%
   ggplot(aes(x = relevel(as.factor(dose_ai.ha), "62.5"), y = prop_YG)) +
   xlab("Dose (g ai/ha)") +
   ggtitle(label = "Yield for each respective propiconazole dose,\nfaceted by number of applications") +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Yield gain proportially\nto the no spray control") +
   facet_grid(cols = vars(total_fungicide))
```

This dose effect should be acknowledged in the meta-analysis.
How many treatments of each dose have been investigated per trial?

```{r table_PropiDoses}
table(as.character(PM_MB_dat[PM_MB_dat$fungicide_ai == "propiconazole",]$trial_ref),
      PM_MB_dat[PM_MB_dat$fungicide_ai == "propiconazole",]$dose_ai.ha)
```

One trial show both doses were applied in the same trial, and mung1112/02.

*****

#### Number of fungicide sprays

Let's look at the frequency of sprays per fungicide.  
```{r Fungicide_x_nSprays}
table(PM_MB_dat$fungicide_ai, PM_MB_dat$total_fungicide)
```

This table shows that the majority of treatments inspected one and two spray management practices.

*****

#### Timing of fungicide sprays  
The timing of fungicide application can be crucial for maximum disease control and yield loss mitigation. 
The experiments all relied on natural infection and thus infection at a standardised crop maturity stage was not possible. 
To best estimate the effect of fungicide application timing we will use the day when the disease was first found in the field as a reference.
The first sign of disease is a useful reference as it can be assessed with little training for the assessor.

First let's make a plot to visualise when one or more spray management regimes were made in relation to the first sign of disease regardless of fungicide active. 
The plot is faceted by number of spray applications.

```{r spray_relative2FS}
PM_MB_dat %>%
   mutate(
      s1_DfromFS = fungicide_application_1 - first_sign_disease,
      s2_DfromFS = fungicide_application_2 - first_sign_disease,
      s3_DfromFS = fungicide_application_3 - first_sign_disease,
      s4_DfromFS = fungicide_application_4 - first_sign_disease,
      s5_DfromFS = fungicide_application_5 - first_sign_disease
   ) %>%
   gather(
      key = spray,
      value = n_days,
      s1_DfromFS,
      s2_DfromFS,
      s3_DfromFS,
      s4_DfromFS,
      s5_DfromFS
   ) %>%
   drop_na(n_days) %>%
   ggplot(aes(x = spray, y = as.numeric(n_days))) +
   geom_violin() +
   scale_x_discrete(labels = c(
      "First spray",
      "Second spray",
      "Third spray",
      "Fourth spray",
      "Fifth spray"
   )) +
   ylab("Fungicide application timing relative to First Sign of Disease") +
   facet_grid(cols = vars(total_fungicide)) +
   coord_flip()
```

The plot shows single fungicide applications are mostly made at zero, first sign of disease in the crop. Some were made earlier and some around 10 days later.  
Experiments that tested two spray treatments also tend to make the first application at first sign of disease, and the second spray between 10 and 25 days later.
Three spray treatments seemed to be mostly used in experiments where preventative sprays as the first fungicide application were made.

*****

### Mungbean cultivars

Australian mungbean varieties have the following resistance to powdery mildew.

   - Berken: Highly susceptible

   - Crystal: Susceptible

   - Jade: Moderately susceptible

Let's view a stacked bar plot of the number of sprays for both demethylation inhibitors, tebuconazole and propiconazole against each cultivar.

```{r Host_genotype}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(host_genotype, fungicide_ai, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Treatments = n) %>%
   ggplot(aes(x = host_genotype, y = Treatments, fill = fungicide_ai)) +
   xlab("Cultivar") +
   ylab("N Trials") +
   ggtitle(label = "Cultivars used in either tebuconazole or propiconazole trials") +
   geom_col() +
   scale_fill_usq(name = "Fungicide AI")
```

#### Genotype yield variability

Does host cultivar significantly impact the yield?
Let's look at the volatility in yields to see if any cultivar yields better or has greater tolerance to powdery mildew.
Below we plot distributions of mean trial yields for each cultivar.
The following plots use data that includes all fungicide types.

First let's look at overall yields for each cultivar, where trial is not a factor, for all treatments (spray and no spray).

```{r yield_vol}
# Outputs the table and graph
source("R/yield_volatility.r") # function to investigate the volatility in yields
# Volatility is the range between the upper and lower 97.5 and 2.5 % quartiles
YV1 <-
   yield_volatility(genotype_by_trial = FALSE, control_only = FALSE)
YV1[[2]] + 
   scale_fill_usq()
YV1[[3]]
```

This plot shows a large amount of yield volatility in all cultivars except Green Diamond.
Green Diamond shows little volatility probably because this cultivar was only used in `r length(unique(PM_MB_dat[PM_MB_dat$host_genotype == "Green Diamond","trial_ref"]))` trial.

Now let's look at overall yields for each cultivar, where trial is not a factor, in only the no spray control treatments.

```{r yieldVol_control}
YV2 <-
   yield_volatility(genotype_by_trial = FALSE, control_only = TRUE)
YV2[[2]] +
   scale_fill_usq()
YV2[[3]]
```

Volatility of the cultivars are less, albeit fairly similar, in control plots when trial is not considered as a factor.

*****

Let's examine the same plots as above but only for the most commonly used trial site, Hermitage.

```{r yieldVol_Herm}
YV1_herm <-
   yield_volatility(
      genotype_by_trial = FALSE,
      control_only = FALSE,
      location = "Hermitage"
   )
YV1_herm[[2]] +
   scale_fill_usq()

YV1_herm[[3]]
```

Jade shows a large amount of volatility, excluding Green Diamond, Crystal shows the next least volatility.

*****

Let's take a closer look at Crystal with spray management treatments and without.

```{r YieldVol_Crystal}
YV1_crystal <-
   yield_volatility(genotype = "Crystal", control_only = FALSE)
YV1_crystal[[2]] +
   scale_fill_usq()

YV1_crystal[[3]]

YV2_crystal <-
   yield_volatility(genotype = "Crystal", control_only = TRUE)
YV2_crystal[[2]] +
   scale_fill_usq()

YV2_crystal[[3]]
```

There is a lot of variation between field trials for Crystal, but little variation within field trials. This indicates that trial should be fit as a random effect when we fit a model to this data.

Let's look at Jade, to see if it differs.

```{r YieldVol_Jade}
YV1_jade <- yield_volatility("Jade",control_only = FALSE)
YV1_jade[[2]] +
   scale_fill_usq()

YV1_jade[[3]]
```

For Jade there is high variation between trials and with-in.
Perhaps indicating lower tolerance. 
Let's look at the same plot with data only from no spray controls to see if the variation is the same.

```{r YieldVol_JadeControl}
YV2_jade <- yield_volatility("Jade", control_only = TRUE)
YV2_jade[[2]] +
   scale_fill_usq()

YV2_jade[[3]]
```

There still remains a higher volatility in Jade and perhaps a greater shift in the probability curve including fungicide treatments versus probability curves including only no spray controls.
Indicating that there might be a difference in powdery mildew tolerance between Jade and Crystal.

Overall there seems to be an influence of cultivar.
However this variation is not as distinct as the variation between trials.

*****

### Row spacing

Some experiments were designed to investigate the effect of row spacing and plant density on powdery mildew disease as well as fungicide timing and efficacy.
The results showed that the row spacing had no statistically significant effect on powdery mildew, but narrower rows in most cases increased yield significantly.
This finding has also been shown by several other researchers' work in Queensland and New South Wales as well.
Eight trials used a row spacing of 0.75&nbsp;meters and tebuconazole as an active ingredient (AI).

```{r row_spacing.m}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   group_by(fungicide_ai, row_spacing, trial_ref) %>%
   summarise() %>%
   count() %>%
   rename(Trials = n) %>%
   ggplot(aes(x = as.factor(row_spacing), y = Trials)) +
   xlab("Row Spacing (m)") +
   ylab("N Trials") +
   ggtitle(label = "Trial row spacing using tebuconazole") +
   geom_col(aes(fill = fungicide_ai),
            position = "dodge") +
   scale_fill_usq(name = "Fungicide AI")
```

Let's plot the row spacing treatments for the response variables yield and disease severity.
The main questions are:

   * Were there any statistical differences for mungbean yield or powdery mildew severity between row spacing treatments?

   * If not, can we pool certain row spacing that have no significant difference?

```{r RS_severity}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   ylab("Final severity rating") +
   xlab("Row spacing (m)") +
   ggtitle("Powdery mildew variation between different row spacing")
```

```{r RS_severity_x_trial}
# Which row spacing leads to the higher disease severity
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   filter(fungicide_ai == "control") %>%
   ggplot(aes(y = PM_final_severity, x = factor(row_spacing))) +
   geom_boxplot(fill = usq_cols("usq charcoal"), alpha = 0.5) +
   facet_grid(cols = vars(location)) +
   ggtitle("Powdery mildew variation between different row spacing") +
   ylab("Final severity rating") +
   xlab("Row spacing (m)")
```

In the Wellcamp site, wider row spacing reduces PM severity, the other sites it seems there was not enough variation to make a distinction between row spacing treatments.

```{r RS_yield}
# Which row spacing leads to the higher yield potential
PM_MB_dat %>%
   filter(year == 2017 |
             year == 2018) %>%
   ggplot(aes(
      y = as.numeric(grain_yield.t.ha.),
      x = as.factor(row_spacing),
      colour = location
   )) +
   geom_jitter(width = 0.1) +
   ggtitle("Grain yield results for different row spacing at three locations") +
   xlab("Row spacing (m)") +
   ylab("Grain yield (t/ha)") +
   scale_colour_usq()
```

The plots seem to imply when yield is limited, presumably by other abiotic factors, there is no effect of row spacing on yield, like in the Hermitage trial.
However, if the average yield is more than approximately 1 t/ha then smaller row spacing has the potential to provide greater yield per hectare.

Overall row spacing may influence both powdery mildew severity and yield.
However, in the studies that were conducted, no interaction was found between row spacing, powdery mildew severity and yield.
That is, the row spacing appeared to increase yield even if the amount of powdery mildew increased and no relationship could be detected between an interaction of the three.

*****

### Yield vs in-season rain

In0-crop rain could be a significant factor which contributes to the end of season grain yield.
To investigate this effect I will use a linear model (GY ~ SR) where seasonal rainfall represents a specific time frame with respect to planting time.
Multiple time frames were tested from 20 days prior to planting, to 100 days after planting.

The multiple time frames were tested by a loop over the dates at the start of the season (to 90 days after planting) to find which date fits the best linear model for grain weight and in-crop rainfall.
The code for this process was moved to be contained within it's own script and can be run as a job as it can take a long time.
To speed up the computation I split the script into four, ran the jobs and bound the data into one frame.
The output of the script is saved with the prefix `lmInSeasonRainfall` followed by the time windows the model is run.
For example `lmInSeasonRainfall_20.40_50.80.csv` holds the model results for start days between 20 and 40 days after planting to end window dates between 50 and 80 days after planting.

This tests when rainfall during the cropping season could be best associated with yield

```{r rainfall_sum_Sstart}
# The four data files were created from four scripts run as four jobs in RStudio
# when the four jobs are finished they can be imported and combined into one data.frame
# source(Rainfall_x_cropYield_s-10.09_e30.65.R)
# source(Rainfall_x_cropYield_s-10.09_e65.100.R)
# source(Rainfall_x_cropYield_s10.30_e30.65.R)
# source(Rainfall_x_cropYield_s10.30_e65.10.R)

lm_rain1 <-
   read.csv("cache/lmInSeasonRainfall_-10.09_30.64.csv",
            stringsAsFactors = FALSE)
lm_rain2 <-
   read.csv("cache/lmInSeasonRainfall_-10.09_65.100.csv",
            stringsAsFactors = FALSE)
lm_rain3 <-
   read.csv("cache/lmInSeasonRainfall_10.30_30.64.csv",
            stringsAsFactors = FALSE)
lm_rain4 <-
   read.csv("cache/lmInSeasonRainfall_10.30_65.100.csv",
            stringsAsFactors = FALSE)

lm_rain <- rbind(lm_rain1, lm_rain2, lm_rain3, lm_rain4)
```

Now we have read in the data from the previously computed models let's find the linear model which has:  
   - The lowest P value  
   - the highest adjusted r squared value

```{r WhichLModel}
lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ]
lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ]
lm_rain[which(lm_rain$lm_adj_rsquared == max(lm_rain$lm_adj_rsquared)), ]
```

To visualise the model fits we will use a heat map to describe how the models faired in relation to the rainfall windows.
log of the p value was used to improve the resolution of small very small p values.

```{r rainLMpvalue}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = log(lm_pval))) +
   geom_raster(aes(fill = log(lm_pval)), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_pval == min(lm_rain$lm_pval)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu", direction = -1) 
```

```{r rainLMadjrsquared}
ggplot(lm_rain, aes(x = start_day, y = end_day, z = lm_adj_rsquared)) +
   geom_raster(aes(fill = lm_adj_rsquared), interpolate = TRUE) +
   geom_contour(bins = 15, colour = "white") +
   geom_point(data = lm_rain[which(lm_rain$lm_rsquared == max(lm_rain$lm_rsquared)), ],
              aes(x = start_day, y = end_day),
              colour = " red") +
   scale_fill_distiller(palette = "RdBu") 
```

The approximate period when in-crop rainfall affects the most on the mean harvest is between 14 - 67 days after planting.

*****

### Season range and first incidence

Mungbean is a summer crop and powdery mildew requires cool conditions to establish.
Therefore the disease can be avoided by planting early in summer or spring, avoiding cooler temperatures at the end of some seasons.
Can we visualise when PM is most likely to establish over the calendar year?

I'll build a plot of horizontal lines indicating the season length and use a dot on the line to indicating the day of the year first sign was observed for that trial.

```{r first_incidence}
# Create the data frame for the plot
season_dates <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(
      Planting_date = unique(planting_date),
      Harvest_date = unique(harvest_date),
      First_sign_PM = unique(first_sign_disease),
      season_length = as.Date(unique(harvest_date)) - as.Date(unique(planting_date)),
      cropping_season = "Actual"
   )

# For some trials where planting date and Harvest date is missing let's estimate 
#     the season length by using the mean of all season lengths
season_dates[is.na(season_dates$season_length),"cropping_season"] <- "Estimated"

season_dates[is.na(season_dates$Planting_date),"Planting_date"] <- 
   season_dates[is.na(season_dates$Planting_date),"Harvest_date"] - 
   as.numeric(mean(season_dates$season_length, na.rm = TRUE))

season_dates[is.na(season_dates$Harvest_date),"Harvest_date"] <- 
   season_dates[is.na(season_dates$Harvest_date),"Planting_date"] + 
   as.numeric(mean(season_dates$season_length, na.rm = TRUE))

season_dates[is.na(season_dates$season_length),"season_length"] <- 
   mean(season_dates$season_length, na.rm = TRUE)

season_dates$Planting_day <- yday(season_dates$Planting_date)
season_dates$Disease_day <- yday(season_dates$First_sign_PM)

season_dates[c(1:2,6), "Planting_day"] <- 1

season_dates$Harvest_day <-
   season_dates$Planting_day + season_dates$season_length

first_day_month <- c(0, 31, 60, 91, 121, 152, 182, 212, 244, 274) + 1
axis_labels_date <-
   c(
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[1], "%b-%d"),
      # Jan
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[2], "%b-%d"),
      # Feb
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[3], "%b-%d"),
      # March
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[4], "%b-%d"),
      # April
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[5], "%b-%d"),
      # May
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[6], "%b-%d"),
      # June
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[7], "%b-%d"),
      # July
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[8], "%b-%d"),
      # August
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[9], "%b-%d"),
      # Sept
      format(Sys.Date() - yday(Sys.Date()) + first_day_month[10], "%b-%d") # October
   )

# reorder the factors so they will appear in chronological order
# I used First sign because all trials have recorded this
season_dates$trial_ref <-
   factor(season_dates$trial_ref,
          levels = levels(season_dates$trial_ref)[rev(order(season_dates$First_sign_PM))])

# Plot
SeasonPlot <- season_dates %>%
   ggplot() +
   geom_pointrange(aes(
      x = trial_ref,
      y = Disease_day,
      ymin = Planting_day,
      ymax = Harvest_day,
      colour = cropping_season
   )) +
   scale_y_continuous(limits = c(
      0,
      max(
         season_dates$Planting_day +
            season_dates$season_length,
         na.rm = TRUE
      )
   ),
   labels = axis_labels_date,
   breaks = first_day_month) +
   geom_point(aes(x = trial_ref, y = Disease_day), size =2.5) +
   coord_flip() +
   scale_color_manual(values=c("black","grey70"))+
   ggtitle("Season length and when powdery mildew was first spotted\nin each respective trial")

SeasonPlot
```

In most years, regardless of the age of the crop, powdery mildew establishes in March, which is to be expected.
The disease is driven by weather conditions and the plant is vulnerable at any growth stage.

### AUDPC visualise raw data

First let's check we have all the AUDPC observations.

```{r AUDPC_check}
PM_MB_dat %>%
   filter(is.na(AUDPC_m)) %>%
   distinct(trial_ref, location, year,AUDPC_m, first_sign_disease, final_assessment)%>%
   kable(align = "c")
```

The trials `r unique(PM_MB_dat[is.na(PM_MB_dat$AUDPC_m),"trial_ref"]) ` don't have an estimate of AUDPC and need to be filled in. 
The following script estimates AUDPC from the final disease severity rating and the first sign of disease information.

```{r SingleObsAUDPC}
source(here("R/AUDPC_for_single_obs.R"))
write.csv(PM_MB_dat,
          "cache/1911_PM_MB_dat&Ygains.csv", row.names = FALSE)
```

      _How well does 'area under the disease progress curve' (AUDPC) predict grain yield?_

   and

      _Does spray_management scenario mitigate yield loss through reducing AUDPC?_

We would expect to see grain yield decline with increasing AUDPC.
Let's just plot yield vs AUDPC.

```{r Yield_vs_AUDPC}

plot(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat)
abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))

```

There appears to be a weak positive relationship between the two continuous variables.
Does a linear model confirm this?

```{r AUDPClm}
summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))
```

This is not what we would expect and the low R squared value of `r round(summary(lm(grain_yield.t.ha. ~ AUDPC_m, data = PM_MB_dat))$adj.r.squared,3)` perhaps there are additional factors influencing this relationship.

Let's tease out variation due to trial location, season, row spacing, and other seasonal variables are preventing us from seeing the expected relationship.
Let's do the plot again for each trial where there are no missing values.
Note: The plots are ordered by mean yield of the trial.

```{r AUDPCxTrial}
sm2lg <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPC_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPC_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(PM_MB_dat$grain_yield.t.ha., na.rm = TRUE)),
      xlim = c(0, max(PM_MB_dat$AUDPC_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPC_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPC_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

}
```

Based on a GRDC report by Sue Thompson [-@SueThompson2016] we want to know what the yield loss was in on cultivar Jade in the 2016 Trials.
```{r}
dat3 <- PM_MB_dat %>%
   filter(year == "2016",
          host_genotype == "Jade",
          fungicide_ai == "control")%>%
   group_by(location)%>%
   summarise(C_Yield = mean(grain_yield.t.ha., na.rm = TRUE))

dat3$F_Yield <- PM_MB_dat %>%
   filter(year == "2016",
          host_genotype == "Jade",
          fungicide_ai != "control")%>%
   group_by(location)%>%
   summarise(F_Yield = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   pull(F_Yield)

dat3 %>% 
   mutate(diff = F_Yield - C_Yield) %>%
   mutate(percent = (diff/F_Yield)*100)

```

This sequence of graphs show that AUDPC and disease incidence tends not affect yields if the season is a low yielding season, _i.e._ lower than 1 tonne per hectare.
The higher the yield, the greater greater the influence on the disease on the yield especially if the crop is high yielding with AUDPC also being large enough to reduce yield.

Let's have a look again at the same sequence of graphs and analyses for area under the disease progress _stairs_.

```{r Yield_vs_AUDPS}
plot(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_dat)
abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_dat))

summary(lm(grain_yield.t.ha. ~ AUDPS_m, data = PM_MB_dat))

sm2lg <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(TY = mean(grain_yield.t.ha., na.rm = TRUE)) %>%
   arrange(TY)

for (i in sm2lg$trial_ref) {
   dat1 <- PM_MB_dat[PM_MB_dat$trial_ref == i, ]
   if (any(is.na(dat1$grain_yield.t.ha.))) {
      next
   }
   if (any(is.na(dat1$AUDPS_m))) {
      next
   }

   plot(
      grain_yield.t.ha. ~ AUDPS_m,
      data = dat1,
      main = paste(dat1$location[1], (dat1$year[1])),
      ylim = c(0, max(
         PM_MB_dat$grain_yield.t.ha., na.rm = TRUE
      )),
      xlim = c(0, max(PM_MB_dat$AUDPS_m, na.rm = TRUE))
   )
   abline(lm(grain_yield.t.ha. ~ AUDPS_m, data = dat1))
   points(grain_yield.t.ha. ~ AUDPS_m,
          data = dat1[dat1$fungicide_ai == "control",],
          col = "red")

}
```

The plots from the AUPDS is similar to the AUDPC, however according to the linear model the AUDPC is a better, significant fit.

A mixed effect model would be more suitable to investigate the interactive effects of location, season, and trial. 
The mixed effect model will help us decide on the best approach and structure for our meta-analysis model.

### FAO mungbean data

An important consideration for why to carry out work on mungbean diseases is the value the crop provides. 
Lets pull in data from the FAOSTAT website for the area of mungbean crops harvested.

```{r FAO_data_inquiry}
beans <- read.csv(here("data/FAOSTAT_data_7-6-2020.csv"))

glimpse(beans)

beans %>%
   filter(Item == "Beans, dry",
          Element == "Area harvested")%>%
   select(Year,Value)%>%
   ggplot(aes(y = Value, x = Year))+
   geom_point()+
   geom_smooth()

```

```{r}
save(SeasonPlot, file = here("cache/PreliminaryData.Rdata"))
```