### 2021-03-01 03_Yield_MetaAnalysis.Rmd
Rename chunks and filename

---
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r Yield_ME_libraries, message=FALSE, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               kableExtra,
               bomrang,
               RColorBrewer,
               metafor,
               here,
               netmeta,
               multcomp,
               flextable)

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq())
source(here("R/reportP.R"))

# Data
# load("cache/ImportDataAndSelectTrials01.Rdata")

PM_dat_Y <-
   read.csv("cache/PM_yield_clean_data.csv", stringsAsFactors = FALSE)
```

# Grain yield meta-analysis

Let's get started with the analysis by first finding the best model fit that answers our research question.

> When, in relation to PM establishes in the crop, should farmers begin spraying mungbean to mitigate yield loss: before PM establishing, immediately after first sign of PM establishing, or after PM has become established in the crop.

A secondary question to this aim is: given the time at which the first spray occurred does a second spray provide worthwhile yield protection.

To do this, in our model:

   - Grain yield is our response variable (t / ha)

   - Trial, which resolves combinations of categorical variables: year, location, row spacing, fungicide dose and cultivar; is set as a random intercept

### Define Trial
```{r Yield_define_trial}
PM_dat_Y <-
   PM_dat_Y %>%
   mutate(trial = paste(trial_ref,
                        year,
                        location,
                        host_genotype,
                        row_spacing,
                        sep = "_")) 
```

   - We will investigate the effect size of our treatment - spray management,

## metafor analysis

For the first grain yield meta-analysis will be an ARM based model using the `metafor` package [@Viechtbauer2010].

We are using the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.

```{r Yield_Metafor-analysis}
PM_mv <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

summary(PM_mv)
```

### Influence of disease onset

To determine if when in the season the disease occurs influences the efficacy at which fungicide applications protect yield we will create a categorical variable defining `early` or `late` disease onset.

`early` and `late` will be divided by the median days after sowing when signs of the disease were first observed.
```{r median onset}
PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   pull(first_sign_days)%>%
   median()
```
Now lets define these two new variables, `first_sign_days` and `onset` in the data.

```{r define_onset}
PM_dat_Y <- PM_dat_Y %>%
   mutate(first_sign_days = lubridate::ymd(first_sign_disease) -
             lubridate::ymd(planting_date)) %>%
   mutate(onset = case_when(first_sign_days <= 39 ~ "early_onset",
                            first_sign_days > 39 ~ "late_onset"))
```

Inspect a histogram of when first sign occured in the crop
```{r hist_onset}
PM_dat_Y %>%
   ggplot(aes(x = first_sign_days)) +
   geom_histogram() +
   geom_vline(xintercept = 39)
```

To put this onset number into context for plant maturity, 40 days is approximately when flowering begins in the plant.

#### metafor onset analysis

We are adding the `onset` variable as a moderator to test and interaction with the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.

```{r onset_Metafor-analysis}
PM_mv_onset <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management * onset,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

```

```{r}
anova(PM_mv, PM_mv_onset)
```
This comparison tells us the meta-analysis including `onset` explained more residual heterogeneity, as shown by the lower QE value.
However the higher AIC indicates that the model omitting `onset` is a slightly better fit.
Also the chi-squared test uncertain that the `onset` model is significantly better as indicated by the `pval` 0.3258.
The higher log likelihood on the `onset` model does intricate there is some merit to this model and that perhaps with more trials to increase statistical power might provide a useful model.

```{r}
summary(PM_mv_onset)
```

### Influence of disease pressure

To determine if the seasonal disease pressure influences the efficacy at which fungicide applications protect yield we will create a categorical variable defining `high_pressure` or `low_pressure`.

`high_pressure` and `low_pressure` will be divided by the median disease severity in the no spray control treatments at the end of the season.
```{r median_pressure}
PM_dat_Y %>%
   filter(fungicide_ai == "control") %>%
   pull(PM_final_severity)%>%
   median()
```
Now lets define these two new variables, `first_sign_days` and `onset` in the data.

```{r define_disease_pressure}
PM_dat_Y <- PM_dat_Y %>%
   mutate(d_pressure = case_when(PM_final_severity < 8 ~ "low_pressure",
                            PM_final_severity >= 8 ~ "high_pressure"))
```

Inspect a histogram of when first sign occured in the crop
```{r hist_pressure}
PM_dat_Y %>%
   ggplot(aes(x = PM_final_severity)) +
   geom_histogram() +
   geom_vline(xintercept = 8)
```

#### metafor onset analysis

We are adding the `onset` variable as a moderator to test and interaction with the `spray_management` variable as a moderator and an interactive term to the `trial` random variable.

```{r osset_Metafor-analysis}
PM_mv_dp <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management * d_pressure,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

```

```{r}
anova(PM_mv, PM_mv_dp)
```
This comparison tells us the meta-analysis including `d_pressure` explained more residual heterogeneity, as shown by the lower QE value.
However the higher AIC indicates that the model omitting `d_pressure` is a slightly better fit.
Also the chi-squared test marginally significant that the `d_pressure` model is better as indicated by the `pval` 0.0468.
The higher log likelihood on the `d_pressure` model does indicate there is some merit to this model.  
It is worth noting that there were no Late plus schedules that experienced low disease pressure. 
Even though the disease pressure interaction on the moderators was marginally significant (P =0.0468), the division of schedules due to disease pressure (reducing k) makes drawing conclusions risky. More data is needed to explore this possible effect.

```{r}
PM_dat_Y %>%
   group_by(spray_management, d_pressure)%>%
   summarise(n())
```

```{r}
summary(PM_mv_dp)
```

### Test variance-covavriance matrix structure
Let's try the model with an implied simpler variance-covariance matrix "compound symmetry" ("CS") to examine if unstructured matrix is suitable.
```{r}
PM_mv_cs <- rma.mv(
   yi = grain_yield.t.ha.,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list(~ spray_management | trial, ~ 1 | id),
   struct = "CS",
   control = list(optimizer = "optim"),
   data = PM_dat_Y
)

anova(PM_mv, PM_mv_cs)
```

Again `PM_mv` (Full) prevails as the better model (p = 0.0014; AIC = -45.4359) and therefore we should keep the unstructured variance-covariance matrix.

 > Therfore the model `PM_mv`, is a best fit.

## Summarise PM_mv model

The first table in the `PM_mv` output shows tau^2 (between trial variance) of each moderator random effects and the number of occurrences for each treatment in the analysis.
This effectively shows the heterogeneity between experiments for these particular treatments.

The tau^2 also gives us the heterogeneity between trials and indicates `Recommended_plus` showed the highest heterogeneity between experiments followed by `Late_plus`, `Late` then `Recommended`, with `Early` and the no-spray `control` showing the least heterogeneity.

The second table is in two parts (left and right). 
The left part, rho, is the correlation of variation between the specified treatments.
All comparisons were acceptable except for a comparison between `Early` and `Late_plus`, `0.000` rho.
`Early` and `Late_plus` treatments never occurred within the same trial, which is indicated by the right side of the table (hence the warning).

In this result we can see that the `Early` treatment is not significantly different to the intercept, which in this case is the mean of the no spray `control`.
However the other treatments are significantly different from the no-spray control (intercept).
The $Q_M$ [omnibus test](http://www.metafor-project.org/doku.php/tips:models_with_or_without_intercept?s[]=anova) of moderators, shows the moderators significantly influence the model ($Q_M =$ `r PM_mv$QM` $,df =$ `r PM_mv$m`, `r reportP(PM_mv$QMp)`) and we can reject the null hypothesis ($H_0 : \beta_1 = \beta_2 = \beta_3 =\beta_4 = 0$) that there is no difference between the moderators [@Viechtbauer2010].
The analysis shows there is still a significant amount of residual heterogeneity ($Q_E =$ `r PM_mv$QE` $,df=$ `r PM_mv$k - PM_mv$m`, `r reportP(PM_mv$QEp)` ) not captured by the spray management moderator indicating other possible moderators which might influence grain yield.

### Meta-analysis moderator estimates table
Let's present the meta-analysis results for the moderator variables in a table of estimates.

```{r metafor_results_table_yield}
# obtain number of treatments included in each moderator variable
k5 <-
   as.data.frame(table(PM_dat_Y$trial, PM_dat_Y$spray_management)) %>%
   filter(Freq != 0) %>%
   group_by(Var2) %>%
   summarise(n()) %>%
   pull()

# create data.frame
results_mv <- data.frame(
   Moderator = c(
   "Intercept / No Spray control",
   "Early",
   "Late",
   "Late+",
   "Recommended",
   "Recommended+"
   ),
   N = PM_mv$g.levels.k,
   k = k5,
   Effect = round(PM_mv$b, 4),
   se = round(PM_mv$se, 4),
   CI_lower = round(PM_mv$ci.lb, 4),
   CI_upper = round(PM_mv$ci.ub, 4),
   z_val = round(PM_mv$zval, 4),
   p_val = reportP(PM_mv$pval, AsNumeric = FALSE, P_prefix = FALSE)
)

# rename colnames to give table headings
colnames(results_mv)[c(2:5, 7:10)] <-
   c("Treatment",
     "N",
     "k",
     "mu",
     "CI_{L}",
     "CI_{U}",
     "Z",
     "P")

yield_estimates_table <- 
   flextable(results_mv[c(2,5,6,3,4), c(1,3:8,10)]) %>%
   align(j = 3:8, align = "center", part = "all") %>%
   fontsize(size = 8, part = "body") %>%
   fontsize(size = 10, part = "header") %>%
   italic(italic = TRUE, part = "header") %>%
   set_caption("Table 2: Estimated mungbean yield mean difference to the no spray control (intercept) for each spray schedule treatment. Yield estimates (u) were calculated from a network meta-analysis of data obtained from grey literature reports of 'k' field trials undertaken in Eastern Australia. P values indicate statistical significance in comparison to the intercept.") %>%
   autofit() %>%
   footnote(i = 1,
            j = c(2:4,6:8),
            value = as_paragraph(c(
               "number of treatment means categorised to each spray schedule",
               "number of trials with the respective spray schedule",
               "estimated mean yield determined by the meta-analysis",
               "Lower range of the 95% confidence interval",
               "Upper range of the 95% confidence interval",
               "indicates the significance between each respective spray schedule and the no spray control (intercept)"
                                   )),
            ref_symbols = letters[1:6],
            part = "header", inline = TRUE
            )

yield_estimates_table
```

Save table to a word document to use in manuscript
```{r eval=FALSE, include=FALSE}
# Save table as a word document
doc <- officer::read_docx()
doc <- body_add_flextable(doc, value = yield_estimates_table)
print(doc, target = "paper/figures/Table2_yield_estimates.docx")
```

Let's view these comparisons in a plot.

```{r metafor_plot_yied}
results_mv %>%
   filter(Treatment != "control") %>%
   mutate(Treatment = factor(Treatment, levels = rev(
      c( "Early",
         "Recommended",
         "Recommended_plus",
         "Late",
         "Late_plus"
      )
   ))) %>%
   ggplot(aes(Treatment, mu)) +
   geom_hline(
      yintercept = seq(-0.05, 0.3, by = 0.05),
      color = usq_cols("usq charcoal"),
      linetype = 3
   ) +
   geom_point(aes(size = 1 / se), shape = 15) +
   geom_linerange(aes(ymin = `CI_{L}`, ymax = `CI_{U}`)) +
   coord_flip() +
   #ggtitle("Mean spray schedule effect sizes\nand 95% confidence intervals")+
   ylab(expression(paste(
      "Mean yield difference to control (t ha" ^ -1, ")", sep = ""
   ))) +
   scale_x_discrete(
      "Moderator variable",
      labels = c(
         expression("Late"["plus"]),
         expression("Late"["single"]),
         expression("Recommended"["plus"]),
         expression("Recommended"["single"]),
         expression("Early"["single"])
         )
   )
ggsave("paper/figures/Fig2_means_difference.png",
       height = 3,
       dpi = 500)
```

Let's investigate the treatment contrasts

```{r PM_Mv_Contrast_yield}
source("R/simple_summary.R") #function to provide a table that includes the treatment names in the contrasts
summary(glht(PM_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), test = adjusted("none"))
contrast_Ssum <-
   simple_summary(summary(glht(PM_mv, linfct = cbind(
      contrMat(rep(1, 6), type = "Tukey")
   )), test = adjusted("none")))
contrast_Ssum

```

These contrasts can be viewed in a plot
```{r plotContrasts_yield}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PM_mv, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), yaxt = 'n')
axis(
   2,
   at = seq_along(contrast_Ssum$contrast),
   labels = rev(contrast_Ssum$contrast),
   las = 2,
   cex.axis = 0.8
)
```

### Profile plots

Let's inspect the profile plots to ensure the model is not over-fitted.
We expect to see the estimate align with the peak of the curve.
Also that the shape of the line is a curve.
Caution! this will take some time to run.

```{r profile_plots, eval=FALSE, message=FALSE, include=FALSE, results="hide"}
profile(PM_mv, tau2 = 1)
profile(PM_mv, tau2 = 2)
profile(PM_mv, tau2 = 3)
profile(PM_mv, tau2 = 4)
profile(PM_mv, tau2 = 5)
profile(PM_mv, tau2 = 6)
```

****

## netmeta analysis

We can use the `netmeta` package to give a graphical representation of the pairwise comparisons.

Let's analyse the data again using a different statistical approach to see if our outcome with the `metafor` package was robust.
The `netmeta` package uses a frequentist approach to the analysis and focuses on the pairwise comparisons between treatments.

```{r netmeta-analysis}
datPM3 <- PM_dat_Y %>%
   group_by(trial, spray_management, replicates) %>%
   summarize(yi_mean = mean(grain_yield.t.ha.),
             vi_mean = mean(vi)) %>%
   ungroup()

PM_con <- pairwise(
   treat = spray_management,
   n = replicates,
   mean = yi_mean,
   sd = sqrt(vi_mean),
   studlab = trial,
   data = datPM3,
   sm = "MD"
)

net_con <- netmeta(TE,
                   seTE,
                   treat1,
                   treat2,
                   studlab,
                   data = PM_con,
                   sm = "MD")

summary(net_con)
```

Now let's visualise this as a forest plot.

```{r netmeta-forest}
forest(
   net_con,
   reference.group = 1,
   rightcols = c("effect", "ci", "Pscore"),
   rightlabs = "P-Score",
   small.values = "bad"
)
```

The `netmeta` analysis suggests the spray schedule commencing early are no different to any other treatment including the no spray `control`.
It estimates the mean is very similar to the recommended treatments.
The `Recommended_plus` and `Late_plus` treatments show higher mean estimates, however are not significantly different from the `Early` estimate.

```{r netgraphGW}
netgraph(
   net_con,
   plastic = FALSE,
   col = usq_cols("support orange"),
   thickness =  "number.of.studies",
   points = FALSE,
   col.points = usq_cols("usq charcoal"),
   cex.points = 1,
   number.of.studies = TRUE,
   cex.number.of.studies = 1,
   col.number.of.studies = "black",
   bg.number.of.studies = usq_cols("support orange"),
   multiarm = FALSE,
   col.multiarm = usq_cols("support turquiose"),
   pos.number.of.studies = 0.4
)
```

```{r}
netleague(net_con)

decomp.design(net_con)

netsplit(net_con)

nm1 <- netmeasures(net_con)

plot(
   nm1$meanpath,
   nm1$minpar,
   pch = "",
   xlab = "Mean path length",
   ylab = "Minimal parallelism"
)
text(nm1$meanpath, nm1$minpar, names(nm1$meanpath), cex = 0.8)
```

```{r Save_meta_data, eval=FALSE}
save(PM_dat_Y,
     PM_mv,
     contrast_Ssum,
     file = here("cache/Meta-analysisData.Rdata"))
```