### 2021-03-03 index.Rmd
reduce title page

---
title: "A meta-analysis of powdery mildew yield loss mitigation strategies in Australian mungbean"
author: "Paul Melloy, Emerson Del Ponte and Adam H. Sparks"
date: "`r format(Sys.time(), '%d %B, %Y')`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: "Mungbean.bib"
biblio-style: apalike
link-citations: yes
github-repo: OpenPlantPathology/Mungbean_PM
description: "A meta-analysis of fungicide use in mungbean to control for yield loss due to powdery mildew."
---
# Home

This research compendium details the statistical analysis and methods of a meta-analysis of fungicide efficacy trials for controlling powdery mildew in mungbean.
The meta-analysis's main aim is to determine:

> When is the optimum time to spray fungicide to limit the effect of powdery mildew on mungbean yields?

## Contents
This compendium is broken down into four sections:

   1. [Details of the experiments, and application of inclusion criteria in the analysis][Trials considered for inclusion in meta-analysis]  
   2. [Preliminary data exploration][Preliminary Analysis]  
   *including a summary of the cleaned trial data and the treatments*
   3. [Grain yield meta-analysis]  
   4. [Powdery mildew severity Meta-analysis]

These sections detail the methods and process we undertook during this meta-analysis.

The manuscript publishing the results presented in this compendium can be found at ....

Preprint at .........