### 2021-05-27 04_Severity_MetaAnalysis.Rmd
Remove percent efficacy from estimates table

---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Powdery mildew severity Meta-analysis

An additional question to our main aim: does spray schedule have a similar effect on disease severity?

This model will be similar to the grain yield meta-analysis and use the main variables:

   - *Powdery mildew severity* the response variable (1-9 scale)

   - *Trial*, which resolves combinations of categorical random variables: 
      i) year  
      ii) location  
      iii) row spacing  
      iv) cultivar

   - *spray_management* a moderator to evaluate the difference in effect size attributed to fungicide application timing.

   - *id* random variable indicating each treatment is independent

<br>

Load libraries and data  
```{r MEsummary_Libraries2, echo=TRUE,results='hide'}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse,
               kableExtra,
               bomrang,
               RColorBrewer,
               metafor,
               here,
               netmeta,
               multcomp,
               dplyr,
               flextable,
               gam,
               clubSandwich)

source(here("R/reportP.R"))

# Data
PM_dat_D <-
   read.csv("cache/PM_disease_clean_data.csv", stringsAsFactors = FALSE)

# functions
logit <- function(x){
   log((x)/(1 - x))
}

inv_logit <- function(x){
   exp(x)/(1 + exp(x))
}

```

<br>  
<br>

### Define Trial
```{r define_trial}
PM_dat_D <-
   PM_dat_D %>%
   mutate(trial = paste(trial_ref,
                        year,
                        location,
                        host_genotype,
                        row_spacing,
                        sep = "_")) 
```

### Inspect PM severity for normality
Given the meta-analysis assumes powdery mildew severity observations to be a univariate normal distribution $Y_i = N(\mu , \sigma)$, with a mean $\mu$ and standard deviation $\sigma$. We need to inspect severity for a normality.
```{r sev_hist}
hist(PM_dat_D$PM_final_severity)
summary(PM_dat_D$PM_final_severity)
```
We can see here that powdery mildew severity observations are skewed towards the top of the scale, eight and nine in this case.

We can improve the normality by transforming the effect size, disease severity, by squaring.
```{r}
PM_dat_D$yi_square <- PM_dat_D$PM_final_severity ^2
PM_dat_D$vi_square <- (PM_dat_D$vi +1)^2
# Note: Here I added 1 to the variance before it was squared. This is because squaring numbers below one makes the number smaller or remain zero. This does not change the meta-analysis mean estimates but does increase heterogeneity and tao._

hist(PM_dat_D$yi_square)
summary(PM_dat_D$yi_square)
```

Another transformation which might be more suitable for a discrete response is either a `log()` or `logit()` transformation. 
However first we should convert the scale from 1 - 9 to between 0 and 1.

### Transform severity response

Let us also transform the severity onto a percentage scale so we can compare the results of meta-analyses with no transformation, squared severity, and percent severity.

```{r}
# Percent severity conversion
# define ordinal conversion from scale to proportion
s1 <- 0
s2 <- 0.33*0.5
s3 <- 0.5*0.87
s4 <- 0.66*0.75
s5 <- 0.66*0.87
s6 <- 0.66*1
s7 <- 0.75*1
s8 <- 1*0.87
s9 <- 1*1
x <- c(s1, s2, s3, s4, s5, s6, s7, s8, s9)
y <- c(1:9)
a = data.frame(x,y)
plot(a)
```

We can fit a gam curve between the 1 - 9 scale and the new scale between 0 and 1
```{r}
# fit a curve using a gam with spline value 0.2
g1 <- gam(x ~ s(y,spar = 0.1))

# View plot of gam
plot(x = predict(g1, data.frame(y= seq(1,9, by =0.1))), y= seq(1,9, by =0.1))
points(x,y,col = "red", pch = 16)

# predict values for dataset
new_sev <- predict(object =  g1, 
                   newdata = data.frame(y = PM_dat_D[,"PM_final_severity"])
                   )
# Check fit
plot(x = new_sev, y = PM_dat_D[,"PM_final_severity"])
points(x,y,col = "red", pch = 16)
```

```{r}
# one value is a negative, this should be a zero
# new_sev[12] <- 0

# Allocate new variable percent severity
PM_dat_D$yi_percent <- new_sev

# Conversion of variance
# Here I add the variance and mean severity and predict the conversion value
# then subtract the mean severity to obtain the difference as the sample variance
PM_dat_D$vi_percent  <- predict(object =  g1,
                   newdata = data.frame(y = PM_dat_D[,"vi"] + PM_dat_D[,"PM_final_severity"])
                   ) - PM_dat_D$yi_percent

# create log transformed values
PM_dat_D$yi_log <- log(PM_dat_D$yi_percent)
PM_dat_D$vi_log <- (PM_dat_D$vi_percent) / (PM_dat_D$replicates * PM_dat_D$yi_percent^2)

# Correct Inf to 0
PM_dat_D[is.na(PM_dat_D$yi_log),"yi_log"] <- 0
PM_dat_D[PM_dat_D$vi_log == -Inf,"vi_log"] <- 0.001

#logit transformed percentages
PM_dat_D$yi_logit <- logit(PM_dat_D$yi_percent)

# logit transformed variance
PM_dat_D$vi_logit <-
   (PM_dat_D$vi_percent) /
   (PM_dat_D$yi_percent * (1 - PM_dat_D$yi_percent)) ^ 2

# Correct -Inf to zeros
PM_dat_D[is.na(PM_dat_D$yi_logit),"yi_logit"] <- 0

```

<br>  
<br>  
<br>

*****

## Disease severity spray schedule meta-analysis  
First lets look at the results of the meta-analysis without transformation
```{r Metafor-Sev-analysis, cache=TRUE}
PMsev_mv <- rma.mv(
   yi = PM_final_severity,
   vi,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   #control = list(optimizer = "optim"),
   data = PM_dat_D
)
summary(PMsev_mv)
plot(rstandard(PMsev_mv)$z, ylim=c(-3,3), pch=19)
```

This analysis shows the residual are slightly unbalanced and perhaps an effect of time. 
Does squaring the response improve the residuals?

```{r Metafor-Sev-analysis, cache=TRUE}
PMsev_mvSQ <- rma.mv(
   yi = yi_square,
   V = vi_square,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mvSQ)
plot(rstandard(PMsev_mvSQ)$z, ylim=c(-3,3), pch=19)
```
This meta-analysis is more in-line with estimates from the yield meta-analysis. `Early` sprays are not effective at reducing powdery mildew severity and two or more sprays had higher efficacy than single sprays. Recommended sprays were the most effective spray timing at reducing disease severity.

However square severity is not common when evaluating a response that was obtained on a scale.
Log transformed is a more common method of evaluating non-normal disease severity.

Lets assess the log transformed percent severity
```{r Metafor-Sev-analysis_log, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mv_log <- rma.mv(
   yi = yi_log,
   V = yi_log,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mv_log)
plot(rstandard(PMsev_mv_log)$z, ylim=c(-3.2,3), pch=19)
```
The plot of the residuals show the model is not normally distributed.
In addition the estimates don't reflect what can be [observed in the raw data][Disease severity range for each spray schedule].
Log transformed percent severity is therefore no a suitable transformation

Finally lets assess the logit transformed percent severity
```{r Metafor-Sev-analysis_logitP, cache=TRUE}
# remove controls for response ratio meta-analysis
PMsev_mvPer <- rma.mv(
   yi = yi_logit,
   V = yi_logit,
   mods = ~ spray_management,
   method = "ML",
   random = list( ~ spray_management | trial, ~ 1 | id),
   struct = "UN",
   control = list(optimizer = "optim"),
   data = PM_dat_D
)

summary(PMsev_mvPer)

# plot the model residuals
plot(rstandard(PMsev_mvPer)$z, ylim=c(-3,3), pch=19)

```

The estimates and significance from this model better reflect the distribution of raw values. 
In addition the residuals from this model are more evenly distributed.
Let's back transform the estimates.

```{r logit_estimates}
backtransform <- function(m, input_only = FALSE, intercept) {
   if (input_only == FALSE) {
      intercept <- m$b[1]

      dat <- data.frame(
         estimates = inv_logit(c(intercept, m$b[-1] + intercept)),
         StdErr = inv_logit(m$se + intercept) - inv_logit(intercept),
         Zval = c(inv_logit(m$z[1]),
                  inv_logit(m$z[-1] + m$z[1]) - inv_logit(m$z[1])),
         Pval = reportP(m$pval, P_prefix = FALSE),
         CI_lb = c(inv_logit(m$ci.lb[1]), 
                       inv_logit(m$ci.lb[-1] + intercept)),
         CI_ub = c(inv_logit(m$ci.ub[1]), 
                       inv_logit(m$ci.ub[-1] + intercept)), 
         row.names = colnames(m$G)
      )
   }
   if (input_only == TRUE) {
      if (missing(intercept))
         stop("'intercept' argument is empty")
     return( inv_logit(m + intercept) - inv_logit(intercept))
   }
   return(dat)
}

backtransform(PMsev_mvPer)
```

<br>  
<br>  
<br>

### Disease severity moderator contrasts

Calculate the treatment contrasts without back-transformation.

```{r sevContrasts}
source("R/simple_summary.R") #function to provide a table that includes the treatment names in the contrasts

intercept <-PMsev_mvPer$b[1,1]
z <- PMsev_mvPer$zval[1]

contrast_Ssum <-
   simple_summary(summary(glht(PMsev_mvPer, linfct = cbind(
      contrMat(rep(1, 6), type = "Tukey")
   )), test = adjusted("none")))

contrast_Ssum[6:15,] %>%
   flextable()%>%
   autofit()
```

<br>  
<br>  
<br>

```{r plotsevContrasts}
par(mar = c(5, 13, 4, 2) + 0.1)
plot(glht(PMsev_mvPer, linfct = cbind(contrMat(rep(
   1, 6
), type = "Tukey"))), yaxt = 'n')
axis(
   2,
   at = seq_along(contrast_Ssum$contrast),
   labels = rev(contrast_Ssum$contrast),
   las = 2,
   cex.axis = 0.8
)
```

`Recommended_plus` limited disease severity significantly better than any other spray treatment. 
Early sprays were not effective in decreasing disease severity.
Two spray treatments were significantly better than a single application, but the timing of the fungicide application was more critical.

<br>  
<br>  
<br>

### Meta-analysis summary table
```{r metafor_results}
# obtain number of treatments included in each moderator variable
k5 <-
   as.data.frame(table(PM_dat_D$spray_management)) %>%
   filter(Freq != 0) %>%
   pull(Freq)

k6 <-
   as.data.frame(table(PM_dat_D$trial_ref, PM_dat_D$spray_management)) %>%
   filter(Freq != 0) %>%
   group_by(Var2) %>%
   summarise(n()) %>%
   pull()

# create data.frame of percent disease severity estimates
results_mv <- cbind(
   data.frame(
      Moderator = c(
         "Intercept / No Spray control",
         "Early",
         "Late",
         "Late+",
         "Recommended",
         "Recommended+"
      ),
      N = k5,
      k = k6
   ),
   backtransform(PMsev_mvPer))

# Calculate mean differences
Intcpt <- results_mv$estimates[1]
results_mv <- 
   results_mv %>%
   mutate(estimates = estimates - Intcpt,
          CI_lb = CI_lb - Intcpt,
          CI_ub = CI_lb - Intcpt)

# round the numbers
results_mv[,c(4:6,8:9)] <- round(results_mv[,c(4:6,8:9)], 4)

# results_mv <- 
#    results_mv %>%
#    mutate(
#    eff_bar = round(
#       1 - estimates, 
#       4) * 100,
#    eff_CI_high = round(
#       1 - CI_lb
#    , 4) * 100,
#    eff_CI_low = round(
#       1 - CI_ub
#    , 4) * 100
#    
# )

# remove efficacy in control
#results_mv[1,10:12] <- NA

# rename colnames to give table headings
colnames(results_mv)[c(1:12 )] <-
   c("Moderator",
     "N",
     "k",
     "mu",
     "se",
     "Z",
     "P",
     "CI_{L}",
     "CI_{U}")

disease_estimates_table <-
   results_mv[c(2,5,6,3,4), -6] %>%
   flextable() %>%
   align(j = 2:8, align = "center", part = "all") %>%
   fontsize(size = 8, part = "body") %>%
   fontsize(size = 10, part = "header") %>%
   italic(italic = TRUE, part = "header") %>%
   set_caption(
      "Table 3: Estimated powdery mildew severity mean difference for each spray schedule treatment. Estimates were back transformed using a inverse logit from the network meta-analysis of data obtained from grey literature reports of (k) field trials undertaken in Eastern Australia. P values indicate statistical significance in comparison to the intercept."
   ) %>%
   footnote(
      i = 1,
      j = c(2:4,6:8),
      value = as_paragraph(
         c(
            "Number of treatment means categorised to each spray schedule",
            "Number of trials with the respective spray schedule",
            "Estimated mean disease severity",
            "Indicates the significance between each respective spray schedule and the no spray control (intercept)",
            "Lower range of the 95% confidence interval",
            "Upper range of the 95% confidence interval"
         )
      ),
      ref_symbols = letters[c(1:7,4,5)],
      part = "header",
      inline = TRUE
    )%>%
   add_header(Moderator = "",
              N = "",
              k = "",
              mu = "Effect Size",
              se = "Effect Size",
              `CI_{L}` = "Effect Size",
              `CI_{U}` = "Effect Size",
              P  = "Effect Size") %>%
   merge_h( part = "header")%>%
   hline_top(part="all", border = officer::fp_border(color = "black", width = 2))%>%
   width(j = 1, width = 1) %>%
   width(j = 2, width = 0.3) %>%
   width(j = 3, width = 0.3) %>%
   width(j = 4:8, width = 0.6)

disease_estimates_table
```

```{r eval=FALSE, include=FALSE}
# Save table as a word document
doc <- officer::read_docx()
doc <- body_add_flextable(doc, value = disease_estimates_table)
print(doc, target = "paper/figures/Table3_severity_esimates.docx")
```